\documentclass[../../math.tex]{subfiles}
\externaldocument{../../math.tex}
\externaldocument{../basics/foundations}
\externaldocument{../basics/set}
\externaldocument{../basics/elementary_algebra}
\externaldocument{../basics/natural}
\externaldocument{../basics/list}
\externaldocument{rational}
\externaldocument{ordinal}

\begin{document}

\setcounter{chapter}{9}

\chapter{The Cardinals}

The cardinals were already introduced in the previous chapter as part of the
development of the theory of ordinals.  Cardinals represent the size of types,
and are defined here as equivalence classes of types under the existence of
bijections.  If you want to see the definition of the cardinals, the
development of their order, and the definition of the function $\aleph$, see the
previous chapter.  In this chapter, we will describe cardinal arithmetic,
describe infinite cardinals, and prove many useful theorems to make working with
cardinals easier.

\section{Addition of Cardinals}

Remember that for two types $\A$ and $\B$, the type $\A + \B$ is the sum of $\A$
and $\B$, also called their disjoint union.  Given an $a : \A$, we call the
corresponding value in $\A + \B$ $a|$, and given a $b : \B$, we call the
corresponding value in $\A + \B$ $|b$.  To make definitions and proofs simpler
here, we will introduce a new notation.  When working with an arbitrary value in
$\A + \B$, we will often use the notation $a|b$ to represent a value made with
either an $a : \A$ or a $b : \B$.  We can manipulate functions to/from sum types
using this notation, and doing so can compactly represent multiple cases of a
definition or proof at once.

\begin{lemma}
    The sum type operation is well-defined under the equivalence relation $\sim$
    given in Section \ref{card_base}.
\end{lemma}
\begin{proof}
    Let $\A \sim \B$ and $\C \sim \D$.  This means that we have bijections $f :
    \A \to \B$ and $g : \C \to \D$.  We must find a bijection between $\A + \C$
    and $\B + \D$.  Define $h : \A + \C \to \B + \D$ by $h(a|c) = f(a)|g(c)$.
    We can also define $h^{-1} : \B + \D \to \A + \C$ by $h(b|d) =
    f^{-1}(b)|g^{-1}(d)$.  Then
    \[
        h(h^{-1}(b|d))
        = h(f^{-1}(b)|g^{-1}(d))
        = f(f^{-1}(b))|g(g^{-1}(d))
        = b|d
    \]
    and
    \[
        h^{-1}(h(a|b))
        = h^{-1}(f(a)|g(b))
        = f^{-1}(f(a))|g^{-1}(g(b)) =
        a|b,
    \]
    so $h$ is bijective by Theorem \ref{inverse_ex_bijective}.
\end{proof}

\begin{instance}
    Define addition in the cardinals as the binary operation given by Theorem
    \ref{binary_op_ex} and the previous lemma.
\end{instance}

\begin{instance}
    Addition of cardinals is associative.
\end{instance}
\begin{proof}
    Let $\A$, $\B$, and $\C$ be types.  Define the functions
    \[
        f : \A + (\B + \C) \to (\A + \B) + \C
    \]
    and
    \[
        g : (\A + \B) + \C \to \A + (\B + \C)
    \]
    by
    \[
        f(a|(b|c)) = (a|b)|c
    \]
    and
    \[
        g((a|b)|c) = a|(b|c).
    \]
    Then
    \[
        f(g((a|b)|c)) = f(a|(b|c)) = (a|b)|c
    \]
    and
    \[
        g(f(a|(b|c))) = g((a|b)|c) = a|(b|c).
    \]
\end{proof}

\begin{instance}
    Addition of cardinals is commutative.
\end{instance}
\begin{proof}
    Let $\A$ and $\B$ be types.  Define functions
    \[
        f : \A + \B \to \B + \A
    \]
    and
    \[
        g : \B + \A \to \A + \B
    \]
    by
    \[
        f(a|b) = b|a
    \]
    and
    \[
        g(b|a) = a|b.
    \]
    Then
    \[
        f(g(b|a)) = f(a|b) = b|a
    \]
    and
    \[
        g(f(a|b)) = g(b|a) = a|b.
    \]
\end{proof}

\begin{instance}
    Define zero in the cardinals to be $|\E|$, the cardinality of the empty
    type.
\end{instance}

\begin{instance}
    Zero is a left additive identity in the cardinals.
\end{instance}
\begin{proof}
    Let $\A$ be a type.  Define the functions $f : \A + \E \to \A$ and $g : \A
    \to \A + \E$ by $f(_|a) = a$ and $g(a) = |a$.  Note that the first case of
    the definition of $f$ is unnecessary because no values of the empty type
    exist.  Then
    \[
        f(g(a)) = f(|a) = a
    \]
    and
    \[
        g(f(|a)) = g(a) = |a.
    \]
\end{proof}

\begin{instance}
    The order of the cardinals is left additive.
\end{instance}
\begin{proof}
    Let $\A$, $\B$, and $\C$ be types such that there exists an injective
    function $f : \A \to \B$.  Then define a function $g : \C + \A \to \C + \B$
    given by
    \[
        g(c|a) = c|f(a).
    \]
    Now assume that $g(c_1|a_1) = g(c_2|a_2)$.  Then $c_1|f(a_1) = c_2|f(a_2)$.
    If it's the first branch, we have $c_1 = c_2$, and if it's the second
    branch, by the injectivity of $f$, we have $a_1 = a_2$.  Thus, we have
    $c_1|a_1 = c_2|a_2$ as required.
\end{proof}

\begin{instance}
    All cardinals are positive.
\end{instance}
\begin{proof}
    The empty function is vacuously injective to any type.
\end{proof}

\begin{theorem} \label{card_false_0}
    For all types $\A$, if $\A \to \False$, then $0 = |\A|$.
\end{theorem}
\begin{proof}
    The empty function is vacuously injective from $\E$ to $\A$, and since $\A
    \to \False$, the empty function is also vacuously surjective.
\end{proof}

\begin{theorem} \label{card_nz_ex}
    For all types $\U$, if $0 \neq |\U|$, then a value exists in $\U$.
\end{theorem}
\begin{proof}
    This is the contrapositive of the previous theorem.
\end{proof}

\begin{instance}
    The function $|\alpha|$ taking an ordinal to a cardinal is nullitive.
\end{instance}
\begin{proof}
    Since $0$ is defined as the equivalence class of the empty type in both the
    cardinals and the ordinals, this follows by reflexivity.
\end{proof}

\begin{instance}
    The function $|\alpha|$ taking an ordinal to a cardinal is additive.
\end{instance}
\begin{proof}
    Let $\A$ and $\B$ be order types.  We must prove that
    \[
        |[\A] + [\B]| = |[\A]| + |[\B]|.
    \]
    By the definition of cardinal addition, this simplifies to
    \[
        |[\A] + [\B]| = |\A + \B|.
    \]
    The proof will be by Theorem \ref{ord_to_card_eq2}.  Let $f$ be the function
    given by definition \ref{ord_type_init_ord}.  Then define a function $g : \A
    + \B \to \Ord$ given by
    \[
        g(x) =
        \begin{cases}
            f_\A(a) &\text{if $x = a|$} \\
            [\A] + f_\B(b) &\text{if $x = |b$.}
        \end{cases}
    \]
    Before we can use this function to prove the result, we must prove that it
    is always less than $[\A] + [\B]$.  If $x = a|$, then
    \[
        g(x) = f_\A(a) < [\A] \leq [\A] + [\B],
    \]
    and if $x = |b$, then
    \[
        g(x) = [\A] + f_\B(b) < [\A] + [\B].
    \]
    Thus, we now have a function between $\A + \B$ and $\Ord_{[\A] + [\B]}$.  We
    will now prove that the function is bijective.

    For injectivity, let $x$ and $y$ be such that $g(x) = g(y)$.  We have four
    cases for when $x$ and $y$ are values in $\A$ or $\B$.  When $x = a_1|$ and
    $y = a_2|$, me have $f_\A(a_1) = f_\A(a_2)$, and the result follows from the
    injectivity of $f_\A$.  When $x = a_1$ and $y = b_2$, we have $f_\A(a_1) =
    [\A] + f_\B(b_2)$, but $f_\A(a_1) < [\A] \leq [\A] + f_\B(b_2)$, meaning
    that this case is impossible.  The case $x = b_1$ and $y = a_2$ is similar.
    When $x = b_1$ and $y = b_2$, we have $[\A] + f_\B(b_1) = [\A] + f_\B(b_2)$,
    and the result follows from cancelling $[\A]$ and applying the injectivity
    of $f_\B$.  Thus, $g$ is injective.

    For surjectivity, let $\gamma < [\A] + [\B]$.  There will be two cases:
    when $\gamma < [\A]$, and when $\gamma \geq [\A]$.  When $\gamma < \alpha$,
    by the surjectivity of $f_\A$ there exists some $a : \A$ such that
    $f_\A(a) = \gamma$.  This means that $g(a|) = \gamma$ as required.  When
    $\gamma \geq [\A]$, by Theorem \ref{ord_le_ex} we have some $\delta$ such
    that $\gamma = [\A] + \delta$.  Then we have $[\A] + \delta < [\A] + [\B]$,
    and by cancelling $[\A]$ we get $\delta < [\B]$.  Thus, by the surjectivity
    of $f_\B$, there exists some $b : \B$ such that $f_\B(b) = \delta$.  Then
    \[
        g(|b) = [\A] + f_\B(b) = [\A] + \delta = \gamma
    \]
    as required.
\end{proof}

\section{Multiplication of Cardinals}

Remember that for two types $\A$ and $\B$, the type $\A \times \B$ is the
product of $\A$ and $\B$, also called their Cartesian product.  Given an $a :
\A$ and a $b : \B$, we call the corresponding value in $\A \times \B$ $(a, b)$.

\begin{lemma}
    The product type operation is well-defined under the equivalence relation
    $\sim$ given in Section \ref{card_base}.
\end{lemma}
\begin{proof}
    Let $\A \sim \B$ and $\C \sim \D$.  This means that we have bijections $f :
    \A \to \B$ and $g : \C \to \D$.  We must find a bijection between $\A \times
    \C$ and $\B \times \D$.  Define $h : \A \times \C \to \B \times \D$ by $h(a,
    c) = (f(a), g(c))$.  We can also define $h^{-1} : \B \times \D \to \A \times
    \C$ by $h^{-1}(b, d) = (f^{-1}(b), g^{-1}(d))$.  Then
    \[
        h(h^{-1}(b, d))
        = h(f^{-1}(b), g^{-1}(d))
        = (f(f^{-1}(b)), g(g^{-1}(d)))
        = (b, d)
    \]
    and
    \[
        h^{-1}(h(a, c))
        = h^{-1}(f(a), g(c))
        = (f^{-1}(f(a)), g^{-1}(g(c)))
        = (a, c),
    \]
    so $h$ is bijective by Theorem \ref{inverse_ex_bijective}.
\end{proof}

\begin{instance}
    Define multiplication in the cardinals as the binary operation given by
    Theorem \ref{binary_op_ex} and the previous lemma.
\end{instance}

\begin{instance}
    Multiplication of cardinals is multiplicative.
\end{instance}
\begin{proof}
    Let $\A$, $\B$, and $\C$ be types.  Define
    \[
        f : \A \times (\B \times \C) \to (\A \times \B) \times \C
    \]
    and
    \[
        g : (\A \times \B) \times \C \to \A \times (\B \times \C)
    \]
    by
    \[
        f(a, (b, c)) = ((a, b), c)
    \]
    and
    \[
        g((a, b), c) = (a, (b, c)).
    \]
    Then
    \[
        f(g((a, b), c)) = f(a, (b, c)) = ((a, b), c)
    \]
    and
    \[
        g(f(a, (b, c))) = g((a, b), c) = (a, (b, c)).
    \]
\end{proof}

\begin{instance}
    Multiplication of cardinals is commutative.
\end{instance}
\begin{proof}
    Let $\A$ and $\B$ be types.  Define functions
    \[
        f : \A \times \B \to \B \times \A
    \]
    and
    \[
        g : \B \times \A \to \A \times \B
    \]
    given by
    \[
        f(a, b) = (b, a)
    \]
    and
    \[
        g(b, a) = (a, b).
    \]
    Then
    \[
        f(g(b, a)) = f(a, b) = (b, a)
    \]
    and
    \[
        g(f(a, b)) = g(b, a) = (a, b).
    \]
\end{proof}

\begin{instance}
    Zero is an annihilator in the cardinals.
\end{instance}
\begin{proof}
    We must find a bijective function from $\A \times \E$ to $\E$ for any type
    $\A$.  Any value in $\A \times \E$ induces a value in $\E$, which is
    impossible, so $\A \times \E$ can't have any values.  Thus, the empty
    function works, and is surjective because the codomain is empty as well.
\end{proof}

\begin{instance}
    Define one in the cardinals to be $|\S|$, the cardinality of the singleton
    type.
\end{instance}

\begin{instance}
    One is a multiplicative identity in the cardinals.
\end{instance}
\begin{proof}
    Let $\A$ be a type.  Define $f : \A \times \S \to \A$ by $f(a, I) = a$ and
    $g : \A \to \A \times \S$ by $g(a) = (a, I)$.  Then
    \[
        f(g(a)) = f(a, I) = a
    \]
    and
    \[
        g(f(a, I)) = g(a) = (a, I).
    \]
\end{proof}

\begin{instance}
    Cardinal multiplication distributes over addition.
\end{instance}
\begin{proof}
    Let $\A$, $\B$, and $\C$ be types.  Define functions
    \[
        f : \A \times (\B + \C) \to (\A \times \B) + (\A \times \C)
    \]
    and
    \[
        g : (\A \times \B) + (\A \times \C) \to \A \times (\B + \C)
    \]
    by
    \[
        f(a, b|c) = (a, b)|(a, c)
    \]
    and
    \[
        g((a, b)|(a, c)) = (a, b|c).
    \]
    Then
    \[
        f(g((a, b)|(a, c))) = f(a, b|c) = (a, b)|(a, c)
    \]
    and
    \[
        g(f(a, b|c)) = g((a, b)|(a, c)) = (a, b|c).
    \]
\end{proof}

\begin{instance}
    Multiplication of cardinals has the zero property.
\end{instance}
\begin{proof}
    For a contradiction, assume that $0 = |\A \times \B|$ but $0 \neq |\A|$ and
    $0 \neq |\B|$.  By Theorem \ref{card_nz_ex}, we have values $a : \A$ and $b
    : \B$.  Because $0 = |\A \times \B|$, we have a bijection $f : \A \times \B
    \to \E$, meaning that $f(a, b)$ is a value in the empty type, which is
    impossible.
\end{proof}

\begin{instance}
    The order of cardinals is left multiplicative.
\end{instance}
\begin{proof}
    Let $\A$, $\B$, and $\C$ be types such that $|\A| \leq |\B|$.  We must prove
    that $|\C \times \A| \leq |\C \times \B|$.  Because $|\A| \leq |\B|$, there
    exists an injective function $f : \A \to \B$.  Define a new function $g : \C
    \times \A \to \C \times \B$ given by $g(c, a) = (c, f(a))$.  Then if
    \[
        g(c_1, a_1) = g(c_2, a_2),
    \]
    we have
    \[
        (c_1, f(a_1)) = (c_2, f(a_2)),
    \]
    from which we get $c_1 = c_2$ and $f(a_1) = f(a_2)$.  By the injectivity of
    $f$, we get $a_1 = a_2$.  Thus, $(c_1, a_1) = (c_2, a_2)$, showing that $g$
    is injective.
\end{proof}

\begin{instance}
    The function $|\alpha|$ taking an ordinal to a cardinal is unital.
\end{instance}
\begin{proof}
    Since $1$ is defined as the equivalence class of the singleton type in both
    the cardinals and the ordinals, this follows by reflexivity.
\end{proof}

\begin{instance}
    The function $|\alpha|$ taking an ordinal to a cardinal is multiplicative.
\end{instance}
\begin{proof}
    Let $\A$ and $\B$ be order types.  We must prove that
    \[
        |[\A][\B]| = |[\A]|\,|[\B]|.
    \]
    By the definition of cardinal multiplication, this simplifies to
    \[
        |[\A][\B]| = |\A \times \B].
    \]
    The proof will be by Theorem \ref{ord_to_card_eq}.  Let $f$ be the function
    given by Definition \ref{ord_type_init_ord}.  Then define a function $g : \A
    \times \B \to \Ord$ given by
    \[
        g(a, b) = [\A] f_\B(b) + f_\A(a).
    \]
    Before we can use this function to prove the result, we must prove that it
    is always less than $[\A] [\B]$.  Let $a : \A$ and $b : \B$.  Then
    \[
        g(a, b) = [\A] f_\B(b) + f_\A(a) < [\A] f_\B(b) + [\A] = [\A] (f_\B(b) +
        1).
    \]
    We have $f_\B(b) < [\B]$, so $f_\B(b) + 1 \leq [\B]$.  Multiplying by $[\A]$
    on the left we get
    \[
        g(a, b) < [\A] (f_\B(b) + 1) \leq [\A] [\B]
    \]
    as required.  Thus, we now have a function between $\A \times \B$ and
    $\Ord_{[\A] [\B]}$.  We will now prove that the function is bijective.

    For injectivity, let $g(a_1, b_1) = g(a_2, b_2)$.  Given that a value $a_1$
    exists in $\A$, we must have $0 \neq [\A]$.  Expanding $g(a_1, b_1) = g(a_2,
    b_2)$, we have
    \begin{equation} \label{ord_to_card_mult_1}
        [\A] f_\B(b_1) + f_\A(a_1) = [\A] f_\B(b_2) + f_\A(a_2).
    \end{equation}
    We have $f_\A(a_1) < [\A]$, and adding $[\A] f_\B(b_1)$ on the left we get
    \begin{equation} \label{ord_to_card_mult_2}
        [\A] f_\B(b_1) + f_\A(a_1) < [\A] f_\B(b_1) + [\A] = [\A] (f_\B(b_1) +
        1).
    \end{equation}
    Using (\ref{ord_to_card_mult_1}) in this, we get
    \begin{equation} \label{ord_to_card_mult_3}
        [\A] f_\B(b_2) + f_\A(a_2) < [\A] (f_\B(b_1) + 1).
    \end{equation}
    Since $[\A] f_\B(b_2) \leq [\A] f_\B(b_2) + f_\A(a_2)$, we also have
    \begin{equation} \label{ord_to_card_mult_4}
        [\A] f_\B(b_2) < [\A] (f_\B(b_1) + 1),
    \end{equation}
    and we can now cancel $[\A]$ to get
    \begin{equation} \label{ord_to_card_mult_5}
        f_\B(b_2) < f_\B(b_1) + 1.
    \end{equation}
    The same argument works symmetrically to get
    \begin{equation} \label{ord_to_card_mult_6}
        f_\B(b_1) < f_\B(b_2) + 1.
    \end{equation}
    This means that $f_\B(b_2) \leq f_\B(b_1)$ and $f_\B(b_1) \leq f_\B(b_2)$,
    so $f_\B(b_1) = f_\B(b_2)$ by antisymmetry.  By the injectivity of $f_\B$,
    we have $b_1 = b_2$.  Going back to (\ref{ord_to_card_mult_1}), we now have
    \begin{equation} \label{ord_to_card_mult_7}
        [\A] f_\B(b_1) + f_\A(a_1) = [\A] f_\B(b_1) + f_\A(a_2),
    \end{equation}
    and we can now cancel $[\A] f_\B(b_1)$ to get $f_\A(a_1) = f_\A(a_2)$, so
    $a_1 = a_2$ by the injectivity of $f_\A$.  Thus, $(a_1, b_1) = (a_2, b_2)$,
    so $g$ is injective.

    For surjectivity, let $\gamma < [\A][\B]$.  We must have $0 \neq [\A]$,
    since if $0 = [\A]$, we would have $\gamma < 0$.  By Theorem \ref{ord_div},
    there exist ordinals $\varepsilon$ and $\delta$ such that $\gamma = [\A]
    \varepsilon + \delta$ and $\delta < [\A]$.  We have
    \[
        [\A] \varepsilon \leq [\A] \varepsilon + \delta = \gamma < [\A] [\B],
    \]
    and cancelling $[\A]$ we get $\varepsilon < [\B]$.  Because $\delta < [\A]$
    and $\varepsilon < [\B]$, we have values $a : \A$ and $b : \B$ such that
    $f_\A(a) = \delta$ and $f_\B(b) = \varepsilon$.  Then
    \[
        g(a, b) = [\A] f_\B(b) + f_\A(a) = [\A] \varepsilon + \delta = \gamma
    \]
    as required.
\end{proof}

\begin{instance} \label{card_not_trivial}
    The cardinals are not trivial.
\end{instance}
\begin{proof}
    If $0 = 1$, we would have a bijective function $f$ from the singleton type
    to the empty type.  Then $f(I)$ is a value of the singleton type, which is
    impossible.  Thus, we must have $0 \neq 1$.
\end{proof}

\begin{theorem} \label{card_one_pos}
    In the cardinals, $0 < 1$.
\end{theorem}
\begin{proof}
    This follows directly from Theorem \ref{all_pos2} and the previous proof.
\end{proof}

\section{Cardinal Exponentiation}

Given two types $\A$ and $\B$, we can create the type of all functions $\A \to
\B$.  In this sense, we can think of $\to$ as being a binary operation on types.
This leads to cardinal exponentiation.  However, we want to define the
exponential $|\A|^{|\B|}$ to be the cardinality of $\B \to \A$, not $\A \to \B$.
Because of this, we can't quite just say to consider $\to$ as a binary
operation for the purpose of constructing cardinal exponentiation.

\begin{lemma}
    $\to$ is well-defined under the equivalence relation $\sim$ given in Section
    \ref{card_base}, in the reverse sense described above.  More explicitly, if
    $\A \sim \B$ and$\C \sim \D$, then $(\C \to \A) \sim (\D \to \B)$.
\end{lemma}
\begin{proof}
    From $\A \sim \B$ and $\C \sim \D$, we have bijections $f : \A \to \B$ and
    $g : \C \to \D$.  Define
    \[
        h : (\C \to \A) \to (\D \to \B)
    \]
    and
    \[
        h^{-1} : (\D \to \B) \to (\C \to \A)
    \]
    by
    \[
        h(i)(d) = f(i(g^{-1}(d)))
    \]
    and
    \[
        h^{-1}(i)(c) = f^{-1}(i(g(c))).
    \]
    To prove that $h(h^{-1}(i)) = i$ and $h^{-1}(h(i)) = i$, by functional
    extensionality it suffices that
    \[
        h(h^{-1}(i))(d)
        = f(h^{-1}(i)(g^{-1}(d)))
        = f(f^{-1}(i(g(g^{-1}(d)))))
        = i(d)
    \]
    and
    \[
        h^{-1}(h(i))(c)
        = f^{-1}(h(i)(g(c)))
        = f^{-1}(f(i(g^{-1}(g(c)))))
        = i(c).
    \]
\end{proof}

\begin{definition}
    Define cardinal exponentiation $\mu^\kappa$ to be the binary operation given
    by Lemma \ref{binary_op_ex} and the previous lemma.  To be explicit, we have
    $|\A|^{|\B|} = |\B \to \A|$.
\end{definition}

\begin{theorem}
    For all cardinals $\kappa$, we have $\kappa^0 = 1$.
\end{theorem}
\begin{proof}
    Given a type $\A$, we can define a map $f$ from $\E \to \A$ to $\S$ that
    takes all functions to the single value in $\S$, and a map $g$ from $\S$ to
    $\E$ to $\A$ that takes in the value in $\S$ and produces the empty
    function $e$, which is the only possible function.  Then
    \[
        f(g(I)) = f(e) = I
    \]
    and
    \[
        g(f(e)) = g(I) = e.
    \]
\end{proof}

\begin{theorem}
    For all cardinals $\kappa \neq 0$, we have $0 ^ \kappa = 0$.
\end{theorem}
\begin{proof}
    Let $\A$ be a type of cardinality $\kappa$.  Since $0 \neq |\A|$, we have a
    value $a : \A$ by Theorem \ref{card_nz_ex}.  If we had a function $f : \A
    \to \E$, we would have a value in the empty type $f(a)$, which is
    impossible.  Thus, by Theorem \ref{card_false_0}, we have $0^\kappa = 0$.
\end{proof}

\begin{theorem}
    For all cardinals $\kappa$, we have $1^\kappa = 1$.
\end{theorem}
\begin{proof}
    Let $\A$ be a type.  Then define the functions $f : (\A \to \S) \to \S$ and
    $g : \S \to (\A \to \S)$ by $f(h) = I$ and $g(I)(x) = I$.  Then
    \[
        f(g(I)) = f(I)
    \]
    and
    \[
        g(f(h)) = g(I) = h.
    \]
\end{proof}

\begin{theorem}
    For all cardinals $\kappa$, we have $\kappa ^ 1 = \kappa$.
\end{theorem}
\begin{proof}
    Let $\A$ be a type.  Define the functions $f : (\S \to \A) \to \A$ and $g :
    \A \to (\S \to \A)$ by $f(h) = h(I)$ and $g(a)(I) = a$.  Then
    \[
        f(g(a)) = g(a)(I) = a
    \]
    and
    \[
        g(f(h))(I) = g(h(I))(I) = h(I).
    \]
\end{proof}

\begin{theorem}
    For all cardinals $\kappa$, $\mu$, and $\nu$, we have $\kappa ^ {\mu + \nu}
    = \kappa ^ \mu \kappa ^ \nu$.
\end{theorem}
\begin{proof}
    Let $\A$, $\B$, and $\C$ be types.  Then define the functions
    \[
        f : ((\B + \C) \to \A) \to ((\B \to \A) \times (\C \to \A))
    \]
    and
    \[
        g : ((\B \to \A) \times (\C \to \A)) \to ((\B + \C) \to \A)
    \]
    by
    \[
        f(h) = ((\lambda b, h(b|)), (\lambda c, h(|c)))
    \]
    and
    \[
        g(h_1, h_2)(b|c) = h_1(b)|h_2(c).
    \]
    Then
    \[
        f(g(h_1, h_2))
        = ((\lambda b, g(h_1, h_2)(b|)), (\lambda c, g(h_1, h_2)(|c)))
        = (\lambda b, h_1(b), \lambda c, h_2(c))
        = (h_1, h_2)
    \]
    and
    \[
        g(f(h))(b|c)
        = g((\lambda b, h(b|)), (\lambda c, h(|c)))(b|c)
        = h(b|)|h(|c)
        = h(b|c).
    \]
\end{proof}

\begin{theorem}
    For all cardinals $\kappa$, $\mu$ and $\nu$, we have $(\kappa ^ \mu) ^ \nu =
    \kappa ^ {\mu \nu}$.
\end{theorem}
\begin{proof}
    Let $\A$, $\B$, and $\C$ be types.  Define the functions
    \[
        f : (\C \to (\B \to \A)) \to ((\B \times \C) \to \A)
    \]
    and
    \[
        g : ((\B \times \C) \to \A) \to (\C \to (\B \to \A))
    \]
    by
    \[
        f(h)(b, c) = h(c)(b)
    \]
    and
    \[
        g(h)(c)(b) = h(b, c).
    \]
    Then
    \[
        f(g(h))(b, c) = g(h)(c)(b) = h(b, c)
    \]
    and
    \[
        g(f(h))(c)(b) = f(h)(b, c) = h(c)(b).
    \]
\end{proof}

\begin{theorem}
    For all cardinals $\kappa$, $\mu$, and $\nu$, we have $(\kappa \mu)^\nu =
    \kappa ^ \nu \mu ^ \nu$.
\end{theorem}
\begin{proof}
    Let $\A$, $\B$, and $C$ be types.  Then define functions
    \[
        f : (\C \to (\A \times \B)) \to ((\C \to \A) \times (\C \to \B))
    \]
    and
    \[
        g : ((\C \to \A) \times (\C \to \B)) \to (\C \to (\A \times \B))
    \]
    by
    \[
        f(h) = ((\lambda c, P_1(h(c))), (\lambda c, P_2(h(c))))
    \]
    and
    \[
        g(h_1, h_2)(c) = (h_1(c), h_2(c)).
    \]
    Then
    \begin{align*}
        f(g(h_1, h_2))
        &= ((\lambda c, P_1(g(h_1, h_2))(c)), (\lambda c, P_2(g(h_1, h_2))(c)))
            \\
        &= ((\lambda c, P_1(h_1, h_2)(c)), (\lambda c, P_2(h_1, h_2)(c))) \\
        &= ((\lambda c, h_1(c)), (\lambda c, h_2(c))) \\
        &= (h_1, h_2)
    \end{align*}
    and
    \[
        g(f(h))(c)
        = g((\lambda c', P_1(h(c'))), (\lambda c', P_2(h(c'))))(c)
        = (P_1(h(c)), P_2(h(c)))
        = h(c).
    \]
\end{proof}

\begin{theorem}
    $|\Prop| = 2$.
\end{theorem}
\begin{proof}
    Recall that $2$ is equal to $|\S| + |\S|$.  Define functions $f : \Prop \to
    \S + \S$ and $g : \S + \S \to \Prop$ given by
    \[
        f(P) = \begin{cases}
            I| & \text{if $P$} \\
            |I & \text{if $\neg P$}
        \end{cases}
    \]
    and
    \[
        g(I|I) = \True|\False.
    \]
    Then
    \[
        f(g(I|I)) = f(\True|\False) = I|I
    \]
    and
    \[
        g(f(P)) = g(I|I) = P.
    \]
\end{proof}

Notice that because of this, $|\A \to \Prop|$ is equal to $2^{|\A|}$, and more
importantly, by Theorem \ref{power_set_bigger}, we have $|\A| < 2^{|\A|}$ for
all cardinals $\A$.

\section{Relationship Between the Cardinals and Natural Numbers}

Like with the ordinals, we can think of the function $\iN$ taking natural
numbers to cardinals.  We will be proving various things about this function and
relating it to the version of $\iN$ for the ordinals.  As such, for the moment,
we will need a way to differentiate between the $\iN$ of the cardinals and the
$\iN$ of the ordinals.  For now, let $\iN^C$ be the version for the cardinals
and $\iN^O$ be the version for the ordinals.  If just $\iN$ is written, it means
$\iN^C$.

\begin{theorem} \label{ord_to_card_nat}
    For all natural numbers $n$, we have $|\iN^O(n)| = \iN^C(n)$.
\end{theorem}
\begin{proof}
    The proof will be by induction on $n$.  When $n = 0$, the result follows
    from $\iN$ and $|\alpha|$ being nullative.  Now assume that $|\iN^O(n)| =
    \iN^C(n)$.  Then
    \[
        |\iN^O(n + 1)| = |\iN^O(n)| + |\iN^O(1)| = \iN^C(n) + 1 = \iN^C(n + 1),
    \]
    proving the theorem by induction.
\end{proof}

\begin{theorem} \label{from_nat_card}
    For all natural numbers $n$, $\iN^C(n) = |\N_n|$.
\end{theorem}
\begin{proof}
    By the previous theorem, we have $\iN^C(n) = |\iN^O(n)|$.  Then the result
    follows from Theorem \ref{from_nat_ord}.
\end{proof}

\begin{instance}
    The cardinals are characteristic zero.
\end{instance}
\begin{proof}
    If $\iN(S(n)) = 0$, we would have a function $f : \N_{S(n)} \to \E$, and
    $f(n)$ would be a value in the empty type, which is impossible.
\end{proof}

Sadly, even with this instance, the usual proof that $\iN$ is injective doesn't
work here because cardinal addition isn't cancellative.  Instead we have to do a
lot of extra work to prove that it is injective.

\begin{lemma} \label{from_nat_card_inj_lemma}
    For all natural numbers $a$ and $b$, if $\iN(S(a)) \leq \iN(S(b))$, then $a
    \iN(a) \leq \iN(b)$.
\end{lemma}
\begin{proof}
    We have an injective function $f : \N_{S(a)} \to \N_{S(b)}$ and must find an
    injective function from $\N_a$ to $\N_b$.  Define $g : \N_a \to \N$ by
    \[
        g(n) =
        \begin{cases}
            f(a) & \text{if $f(n) = b$} \\
            f(n) & \text{otherwise.}
        \end{cases}
    \]
    We must prove that $g(n) < b$ for all $n < a$.  There are two cases: when
    $f(n) = b$, and when $f(n) \neq b$.  When $f(n) = b$, we have $g(n) = f(a)$.
    We already have $f(a) \leq b$ by the definition of $f$, so we just need to
    prove that $f(a) \neq b$.  If $f(a) = b$, we would have $f(n) = f(a)$, and
    by the injectivity of $f$ we get $n = a$, cnotradicting $n < a$.  Thus, we
    must have $f(a) \neq b$, so $g(n) < b$ in the case $f(n) = b$.  In the case
    $f(n) \neq b$, we have $g(n) = f(n)$.  Since $f(n) \leq b$ by definition and
    $f(n) \neq b$, we have $g(n) < b$.

    Thus, $g$ is a function from $\N_a$ to $\N_b$.  We must prove that $g$ is
    injective.  Let $m < a$ and $n < a$ and $g(m) = g(n)$.  There will be four
    cases, for when $f(m) = b$ and $f(m) \neq b$, and for when $g(m) = b$ and
    $g(m) \neq b$.  When $f(m) = b$ and $f(n) = b$, we have $f(m) = f(n)$, and
    by injectivity of $f$ we have $m = n$ as required.  When $f(m) = b$ and
    $f(n) \neq b$, we have $f(a) = f(n)$, and by injectivity of $f$ we have $a =
    n$, contradicting $n < a$, so this case is impossible.  The case $f(m) \neq
    b$ and $f(n) \neq b$ is similar.  When $f(m) \neq b$ and $f(n) \neq b$, we
    have $f(m) = f(n)$, and by injectivity of $f$ we get $m = n$ as required.
    Thus, $g$ is injective.
\end{proof}

\begin{lemma} \label{from_nat_card_zero_eq}
    For all natural numbers $n$, if $0 = \iN n$, then $0 = n$.
\end{lemma}
\begin{proof}
    Since $\iN n = 0$, we have a bijection $f : \N_n \to \E$.  If $0 \neq n$, we
    have $0 < n$, so $f(0)$ is a value in the empty type, which is impossible.
\end{proof}

\begin{instance}
    $\iN$ is injective.
\end{instance}
\begin{proof}
    Let $\iN(a) = \iN(b)$.  The proof will be byinduction on $a$.  When $a = 0$,
    the result follows by Lemma \ref{from_nat_card_zero_eq}.  Now assume that
    for all $b'$ such that $\iN(a) = \iN(b')$, we have $a = b'$.  We have
    $\iN(S(a)) = \iN(b)$ and must prove $S(a) = b$.  if $b = 0$, we would have
    $S(a) = 0$ by Lemma \ref{from_nat_card_zero_eq}, which is impossible.  So
    there exists a $b'$ such that $b = S(b')$.  Then we have $\iN(S(a)) =
    \iN(S(b'))$, and by antisymmetry and Lemma \ref{from_nat_card_inj_lemma}, we
    have $\iN(a) = \iN(b')$.  Thus, by the inductive hypothesis, we have $a =
    b'$, so $S(a) = S(b') = b$ as required.
\end{proof}

\begin{theorem} \label{card_to_initial_ord_nat}
    For all natural numbers $n$, $\lfloor \iN^C(n) \rfloor = \iN^O(n)$.
\end{theorem}
\begin{proof}
    By Theorem \ref{ord_to_card_nat}, we have $\lfloor \iN^C(n) \rfloor =
    \lfloor | \iN^O(n) | \rfloor$.  To prove that $\lfloor | \iN^O(n) |
    \rfloor$, we can use Theorem \ref{card_to_initial_ord_other_eq}.  We must
    prove that for all ordinals $\beta < \iN^O(n)$, we have $|\beta| <
    |\iN^O(n)|$.  Because $\beta < \iN^O(n)$, we have $\beta < \omega$, so there
    exists a natural number $m$ such that $\beta = \iN^O(m)$.  Thus, we have
    $\iN^O(m) < \iN^O(n)$, so $m < n$.  We now must prove that $|\iN^O(m)| <
    |\iN^O(n)|$.  We have $|\iN^O(m)| \leq |\iN^O(n)|$ from Theorem
    \ref{ord_to_card_le}.  We have $|\iN^O(m)| \neq |\iN^O(n)|$ from Theorem
    \ref{ord_to_card_nat} and the fact that $\iN^C$ is injective.
\end{proof}

\begin{instance}
    $\iN^C$ is orderly.
\end{instance}
\begin{proof}
    Let $a \leq b$.  Then $\iN^O(a) \leq \iN^O(b)$ by $\iN^O$ being orderly.
    Then $\lfloor \iN^C(a) \rfloor \leq \lfloor \iN^C(b) \rfloor$ by Theorem
    \ref{card_to_initial_ord_nat}.  Then $\iN^C(a) \leq \iN^C(b)$ by $\lfloor
    \mu \rfloor$ being equivalently orderly.
\end{proof}

Now that we have proven these basic properties of $\iN^C$, we don't need to
consider it to be any different from $\iN^O$, and $\iN$ will be dropped entirely
for cardinals/ordinals now.

\begin{theorem} \label{card_omega}
    $|\omega| = |\N|$.
\end{theorem}
\begin{proof}
    This is true by definition.
\end{proof}

\begin{theorem} \label{nat_lt_card}
    For all natural numbers $n$, we have $n < |\N|$.
\end{theorem}
\begin{proof}
    Assume that $|\N| \leq n$.  This means that there exists an injective
    function from $\N$ to $\N_n$.  Thus, $\N$ is simply finite, contradicting
    Theorem \ref{nat_not_finite}.  So we must have $n < |\N|$.
\end{proof}

\begin{theorem} \label{card_omega_initial}
    $\lfloor | \N | \rfloor = \omega$.
\end{theorem}
\begin{proof}
    The proof will be by Theorem \ref{card_to_initial_ord_other_eq}.  We must
    prove that for all ordinals $\beta < \omega$, we have $|\beta| < |\omega|$.
    Because $\beta \omega$, we know that $\beta$ is a natural number.  Then the
    result follows from Theorem \ref{nat_lt_card}.
\end{proof}

\begin{theorem} \label{card_lt_nat}
    For all cardinals $\kappa < |\N|$, there exists a natural number $n =
    \kappa$.
\end{theorem}
\begin{proof}
    Because $\kappa < |\N|$, we have $\lfloor \kappa \rfloor < \lfloor |\N|
    \rfloor = \omega$, so by Theorem \ref{ord_lt_omega} there exists a natural
    number $n = \lfloor \kappa \rfloor$.  Then by Theorem
    \ref{card_to_initial_ord_to_card_eq}, we have $n = \kappa$.
\end{proof}

\begin{theorem} \label{card_lt_one_eq}
    For all cardinals $\kappa$, if $\kappa < 1$, then $0 = \kappa$.
\end{theorem}
\begin{proof}
    Since $\kappa < 1$, we have $\kappa < |\N|$, so $\kappa$ is a natural
    number.  Then the theorem follows from the corresponding result in the
    natural numbers.
\end{proof}

\begin{theorem} \label{card_pos_one}
    For all cardinals $\kappa$, we have $0 \neq \kappa$ if and only if $1 \leq
    \kappa$.
\end{theorem}
\begin{proof}
    If $0 \neq \kappa$ and $\kappa < 1$, we have $0 = \kappa$ by the previous
    theorem, contradicting $0 \neq \kappa$.  Thus, when $0 \neq \kappa$, we have
    $1 \leq \kappa$.  Now when $1 \leq \kappa$, if we have $0 = \kappa$, we
    would have $1 \leq 0$, which is impossible.  Thus, if $1 \leq \kappa$, we
    have $0 \neq \kappa$.
\end{proof}

\begin{theorem} \label{aleph'_nat}
    For all natural numbers $n$, we have $\aleph'_n = n$.
\end{theorem}
\begin{proof}
    The proof will be by strong induction on $n$.  The inductive hypothesis
    states that for all $m < n$, we have $\aleph'_m = m$.  We will prove that
    $\aleph'_n = n$ by antisymmetry.  To prove that $\aleph'_n \leq n$, we will
    use Theorem \ref{aleph'_least}.  Let $m < n$.  Then by the inductive
    hypothesis, $\aleph'_m = m < n$, so we have $\aleph'_n \leq n$.  To prove
    that $n \leq \aleph'_n$, assume for a contradiction that $\aleph'_n < n$.
    Then there exists a natural number $m = \aleph'_n$, so $m < n$.  By the
    inductive hypothesis we have $\aleph'_m = m$, so $\aleph'_m = \aleph'_n$.
    Because $\aleph'$ is injective, we have $m = n$, contradicting $m < n$.
    Thus, $n \leq \aleph'_n$, so by antisymmetry we have $\aleph'_n = n$.
\end{proof}

\begin{theorem} \label{aleph_0}
    $\aleph_0 = |\N|$.
\end{theorem}
\begin{proof}
    We must prove that $\aleph'_\omega = |\N|$.  We will prove it by
    antisymmetry.  To prove that $\aleph'_\omega \leq |\N|$, we will use Theorem
    \ref{aleph'_least}.  Let $m < \omega$.  By Theorem \ref{ord_lt_omega}, $m$
    is a natural number.  Then $\aleph'_m = m < |\N|$, showing that
    $\aleph'_\omega \leq |\N|$.  To prove that $|\N| \leq \aleph'_\omega$,
    assume for a contradiction that $\aleph'_\omega < |\N|$.  Then there exists
    a natural number $m$ such that $m = \aleph'_\omega$  By Theorem
    \ref{aleph'_nat}, we have $\aleph'_m = \aleph'_\omega$, so $m = \omega$,
    which is impossible.  Thus, $|\N| \leq \aleph'_\omega$, so $\aleph_0 = |\N|$
    by antisymmetry.
\end{proof}

\section{Infinite and Finite Cardinals}

\begin{definition}
    A cardinal $\kappa$ is finite if $\kappa < |\N|$.
\end{definition}

\begin{definition}
    A cardinal $\kappa$ is infinite if $|\N| \leq \kappa$.
\end{definition}

Because the cardinals are totally ordered, every cardinal is either finite or
infinite, and can't be both.

\begin{theorem} \label{nat_is_finite}
    Every natural number is finite.
\end{theorem}
\begin{proof}
    This is just restating Theorem \ref{nat_is_finite}.
\end{proof}

\begin{theorem} \label{fin_nat_ex}
    For all finite cardinals $\kappa$, there exists an $n = \kappa$,
\end{theorem}
\begin{proof}
    This is just restating Theorem \ref{card_lt_nat}.
\end{proof}

\begin{theorem} \label{greater_all_nat_inf}
    For all cardinals $\kappa$, if $\kappa > n$ for all natural numbers $n$,
    then $\kappa$ is infinite.
\end{theorem}
\begin{proof}
    Assume that $\kappa$ is finite.  Then by Theorem \ref{fin_nat_ex}, there
    exists some natural number $n$ with $n = \kappa$.  Then $n > n$, a
    contradiction.  Thus, $\kappa$ is infinite.
\end{proof}

\begin{theorem} \label{inf_not_nat}
    For all infinite cardinals $\kappa$, $\kappa \neq n$ for all natural numbers
    $n$.
\end{theorem}
\begin{proof}
    Direct from Theorem \ref{nat_is_finite}.
\end{proof}

\begin{theorem} \label{simple_finite_finite}
    For all types $\U$, $\U$ is simply finite if and only if $|\U|$ is finite.
\end{theorem}
\begin{proof}
    Assume that $\U$ is simply finite.  Then there exists an injective function
    from $\U$ to $\T(n)$ for some natural number $n$, so $|\U| \leq n$.  Since
    $n < |\N|$, $|\U|$ is finite.

    Now assume that $|\U|$ is finite.  This means that $|\U| = n$ for some
    natural number $n$.  Thus, there exists a bijective function $f$ from $\U$
    to $\T(n)$.  Because $f$ is injective, $\U$ is simply finite.
\end{proof}

Thus, all of the theorems proven for simply finite types apply to finite types
too.

\begin{theorem} \label{infinite_nz}
    For all infinite cardinals $\kappa$, $0 \neq \kappa$.
\end{theorem}
\begin{proof}
    Special case of Theorem \ref{inf_not_nat}.
\end{proof}

\begin{theorem} \label{infinite_ex}
    For all infinite types $\U$, there exists a value in $\U$.
\end{theorem}
\begin{proof}
    Direct from Theorem \ref{card_nz_ex} and the previous theorem.
\end{proof}

\begin{theorem} \label{all_greater_inf}
    For all ordered types $\U$, if there exists a value $a : \U$ and if for all
    $b : \U$, there exists a $c > b$, then $\U$ is infinite.
\end{theorem}
\begin{proof}
    Define a function $f : \N \to \U$ recursively as
    \[
        f(n) = \begin{cases}
            a & \text{if $n = 0$} \\
            \text{a value greater than $f(n-1)$} &\text{otherwise,}
        \end{cases}
    \]
    where the recursive case always exists by assumption.  Now let $f(m) =
    f(n)$.  We must prove that $m = n$.  For a contradiction, assume that $m
    \neq n$, and without loss of generality assume that $m < n$.  By Theorem
    \ref{nat_lt_ex}, we have some $c : \N$ such that $n = m + S(c)$.  So we now
    must prove that $f(m) < f(m + S(c))$, which we will do by induction on $c$.

    When $c = 0$, we have $f(m) < f(m + 1)$ by the definition of $f$, so the
    base case is true.  Now assume that $f(m) < f(m + S(c))$.  Then
    \[
        f(m) < f(m + S(c)) = f(S(m + S(c))) = f(m + S(S(c))),
    \]
    showing that the inductive case is true.  Thus, by induction, we have $f(m)
    < f(m + S(c)) = f(n)$.  However, this contradicts $f(m) = f(n)$, so our
    original assumption that $m < n$ must be false.  Thus, $m = n$, proving that
    $f$ is injective, so $\U$ is infinite.
\end{proof}

\section{Arithmetic of Infinite Cardinals}

It turns out that arithmetic of infinite cardinals is trivial in many aspects.
However, to get to that point, we will have to prove several other things first.

\begin{lemma} \label{card_inf_plus_one}
    For all infinite cardinals $\mu$, $\mu + 1 = \mu$.
\end{lemma}
\begin{proof}
    Let $\U$ be a type with cardinality $\mu$.  Because $\U$ is infinite, there
    exists an injective function $f : \N \to \U$.  Now define a functions $g :
    \U + \S \to \U$ and $h : \U \to \U + \S$ by
    \[
        g(x) = \begin{cases}
            f(S(n)) & \text{if there exists an $n$ with $f(n)| = x$} \\
            a       & \text{if no such $n$ exists and $a| = x$} \\
            f(0)    & \text{if $|I = x$}
        \end{cases}
    \]
    and
    \[
        h(a) = \begin{cases}
            |I    & \text{if $a = f(0)$} \\
            f(n)| & \text{if there exists an $n$ with $f(S(n)) = a$} \\
            a|    & \text{otherwise.}
        \end{cases}
    \]

    Now consider an arbitrary value $a : \U$.  If $a = f(0)$,
    \[
        g(h(a)) = g(|I) = f(0) = a.
    \]
    If there exists an $n$ with $f(S(n)) = a$,
    \[
        g(h(a)) = g(f(n)|) = f(S(n)) = a.
    \]
    Otherwise,
    \[
        g(h(a)) = g(a|) = a.
    \]
    In all cases, $g(h(a)) = a$.

    Now consider an arbitrary value $x : \U + \S$.  If there exists an $n$ with
    $f(n)| = x$,
    \[
        h(g(x)) = h(f(S(n))|) = f(n)| = x.
    \]
    If no such $n$ exists and $a| = x$,
    \[
        h(g(x)) = h(a) = a| = x.
    \]
    Otherwise, $|I = x$, so
    \[
        h(g(x)) = h(f(0)) = |I = x.
    \]
    In all cases, $h(g(x)) = x$.

    Thus, by Theorem \ref{inverse_ex_bijective}, $g$ is bijective, so $\mu + 1 =
    \mu$.
\end{proof}

\begin{lemma} \label{card_inf_minus_one_inf}
    For all cardinals $\mu$, if $\mu + 1$ is infinite, then $\mu$ is infinite.
\end{lemma}
\begin{proof}
    Assume that $\mu$ is not infinite.  Then $\mu$ is a natural number, meaning
    that $\mu + 1$ is a natural number too, contradicting $\mu + 1$ being
    infinite.  Thus, $\mu$ is infinite.
\end{proof}

\begin{lemma} \label{card_inf_remove_one}
    For all types $\U$ and values $a : \U$, if $\U$ is infinite, then
    \[
        |\U| = |\{x : \U, x \neq a\}|.
    \]
\end{lemma}
\begin{proof}
    Denote the set $\{x : \U, x \neq a\}$ as $A$.  Then $|A| + 1 = |\U|$ by
    bringing everything in $A$ to its value in $\U$, and then taking the
    singleton value to $a$.  Then by Lemma \ref{card_inf_minus_one_inf}, $|A|$
    is infinite too, so by Lemma \ref{card_inf_plus_one}, $|A| + 1 = |A|$, so
    $|\U| = |A|$.
\end{proof}

We can now start moving towards the main result.  It involves many substeps and
lemmas, all of which work from an infinite order type $\A$ as its starting
point.  So from this point up until Lemma \ref{ord_mult_idemp}, let $\A$ be an
arbitrary infinite order type.

\begin{instance}
    Order the type $\A \times \A$ by saying that $(a_1, a_2) \leq (b_1, b_2)$ if
    \begin{align*}
        &(\max(a_1, a_2) < \max(b_1, b_2)) \text{ or} \\
        &(\max(a_1, a_2) = \max(b_1, b_2) \text{ and } a_1 < b_1) \text{ or} \\
        &(\max(a_1, a_2) = \max(b_1, b_2) \text{ and } a_1 = b_1 \text{ and }
            a_2 \leq b_2).
    \end{align*}
\end{instance}

\begin{instance}
    The order on $\A \times \A$ is antisymmetric.
\end{instance}
\begin{proof}
    Let $(a_1, a_2) \leq (b_1, b_2)$ and $(b_1, b_2) \leq (a_1, a_2)$.  Since
    each inequality has three cases, it would appear that there are nine cases.
    However, in all but one of those cases, we have either two things strictly
    less than each other, which is impossible, or two things that are equal but
    also with one strictly less than the other, which is also impossible.  The
    only possible cases is when $\max(a_1, a_2) = \max(b_1, b_2)$, $a_1 = b_1$,
    $a_2 \leq b_2$, and $b_2 \leq a_2$.  Then by antisymmetry we have $a_2 =
    b_2$, and since $a_1 = b_1$ and $a_2 = b_2$, we have $(a_1, a_2) = (b_1,
    b_2)$.
\end{proof}

\begin{instance}
    The order on $\A \times \A$ is a well order.
\end{instance}
\begin{proof}
    Let $S$ be a nonempty set in $\A \times \A$.  Define a set $S'$ by
    \[
        S' = \{a : \A \mid
            \exists a_1 a_2, (a_1, a_2) \in S \wedge a = \max(a_1, a_2)\}.
    \]
    Because $S$ is nonempty, this has a least element $d$.  Now define a set
    $S''$ by
    \[
        S'' = \{a_1 : \A \mid
            \exists a_2, (a_1, a_2) \in S \wedge d = \max(a_1, a_2)\}.
    \]
    Because $d \in S'$, $S''$ is nonempty, so it has a least element $x$.  Now
    define a set $S'''$ by
    \[
        S''' = \{a_2 : \A \mid (x, a_2) \in S \wedge d = \max(x, a_2)\}.
    \]
    Then because $x \in S''$, $S'''$ is nonempty, so it has a least element $y$.
    Note that $d = \max(x, y)$.  We will now prove that $(x, y)$ is the least
    element of $S$.  First, $(x, y) \in S$ by definition, so we only need to
    prove that it is the least element of $S$.  So let $(a_1, a_2) \in S$.  We
    must prove that $(x, y) \leq (a_1, a_2)$.

    If $\max(x, y) = \max(a_1, a_2)$ and $x = a_1$, then $a_2 \in S'''$, so $y
    \leq a_2$ by the minimality of $y$.  Thus, $(x, y) \leq (a_1, a_2)$.

    If $\max(x, y) = \max(a_1, a_2)$ and $x \neq a_1$, then $a_1 \in S''$, so $x
    \leq a_1$ by the minimality of $x$.  Thus, $x < a_1$, so $(x, y) \leq (a_1,
    a_2)$.

    If $\max(x, y) \neq \max(a_1, a_2)$, then $\max(a_1, a_2)$ is in $S'$, so
    $\max(x, y) = d \leq \max(a_1, a_2)$ by the minimality of $d$.  Thus,
    $\max(x, y) < \max(a_1, a_2)$, so $(x, y) \leq (a_1, a_2)$.

    In all cases $(x, y) \leq (a_1, a_2)$, showing that $(x, y)$ is the least
    element of $S$, proving that $\A \times \A$ is well ordered.
\end{proof}

We will now be working up to the main lemma that $|\A| = |\A \times \A|$.  This
will be done by transfinite induction.  The inductive hypothesis states that for
all infinite order types $\B$ with $[\B] < [\A]$, we have $|\B| = |\B \times
\B|$.  So until Lemma \ref{ord_mult_idemp}, this inductive hypothesis will be
valid.

\begin{lemma} \label{not_card_idemp}
    If there exists an order type $\B$ with $[\B] < [\A]$ and $|\B| = |\A|$,
    then $|\A| = |\A \times \A|$.
\end{lemma}
\begin{proof}
    Since $|\A| = |\B|$ and $|\B| = |\B \times \B|$ by the inductive hypothesis,
    we have $|\A| = |\A \times \A|$.
\end{proof}

We will now consider the case when the assumption of the previous lemma is not
true, that is, when for all ordinals $\beta < [\A]$, we have $|\beta| \neq
|\A|$.  This assumption will hold until Lemma \ref{ord_mult_idemp}.  Call this
assumption the cardinality assumption.

\begin{lemma} \label{card_mult_idemp_gt_ex}
    There exists a $b > a$ for all $a : \A$.
\end{lemma}
\begin{proof}
    Assume otherwise, that is, there exists an $a : \A$ with $b \leq a$ for all
    $b : \A$.  Then with $f$ being the function given by Definition
    \ref{ord_type_init_ord}, we have $f_\A(a) < [\A]$, and if we can prove that
    $|f_\A(a)| = |\A|$, we will have a contradiction with the cardinality
    assumption.  By Theorem \ref{ord_to_card_init}, we have $|f_\A(a)| =
    |\A_a|$, so it suffices to prove that $|\A_a| = |\A|$.  Because $a$ is the
    maximum element of $\A$, $\A_a$ is just the set $\{x : \A \mid x \neq a\}$.
    Then $|\A_a| = |\A|$ by Lemma \ref{card_inf_remove_one}.
\end{proof}

\begin{lemma} \label{card_mult_idemp_l_leq}
    For all $l_1$, $l_2$, and $l$ in $\A$, if $\max(l_1, l_2) < l$, then
    \[
        |(\A \times \A)_{(l_1, l_2)} \leq
        |\A_l| \, |\A_l|.
    \]
\end{lemma}
\begin{proof}
    First, we will prove that for all $(d_1, d_2) < (l_1, l_2)$, we have $d_1 <
    l$ and $d_2 < l$.  When $\max(d_1, d_2) < \max(l_1, l_2) < l$, we have $d_1
    < l$ and $d_2 < l$.  When $\max(d_1, d_2) = \max(l_1, l_2)$, we also have
    $d_1 < l$ and $d_2 < l$.  In both cases, we have $d_1 < l$ and $d_2 < l$.
    Then the function taking $(d_1, d_2)$ in $(\A \times \A)_{(l_1, l_2)}$ to
    $(d_1, d_2)$ in $\A_l \times \A_l$ is an injective function.
\end{proof}

\begin{lemma} \label{card_mult_idemp_l_lt}
    For all $l_1$ and $l_2$ in $\A$, with $f$ the function given by Definition
    \ref{ord_type_init_ord}, we have $f_{\A \times \A}(l_1, l_2) < [\A]$.
\end{lemma}
\begin{proof}
    First, by Lemma \ref{card_mult_idemp_gt_ex}, we have some $l$ with $l_1 < l$
    and $l_2 < l$.  By Theorems \ref{ord_to_card_lt} and \ref{ord_to_card_init},
    it suffices to prove that $|(\A \times \A)_{(l_1, l_2)}| < |\A|$.  By
    transitivity and Lemma \ref{card_mult_idemp_l_leq}, it suffices to prove
    that $|\A_l| \, |\A_l| < |\A|$.  There will be two cases: when $|\A_l|$ is
    infinite, and when it is finite.

    When $|\A_l|$ is infinite, by the inductive hypothesis we have $|\A_l| =
    |\A_l \times \A_l|$.  Then we must prove that $|\A_l| < |\A|$.  By Theorem
    \ref{ord_to_card_le}, we already have $|\A_l| \leq |\A|$, so we just need to
    prove that $|\A_l| \neq |\A|$.  This follows directly from the cardinality
    assumption.

    When $|\A_l|$ is finite, $|\A_l| \, |\A_l|$ is finite as well, and since
    $|\A|$ is infinite, we have $|\A_l| \, |\A_l| < |\A|$.
\end{proof}

\begin{lemma} \label{card_mult_idemp_A_leq}
    $[\A \times \A] \leq [\A]$.
\end{lemma}
\begin{proof}
    Assume that $[\A] < [\A \times \A]$.  Then because the ordinal initial
    segment function $f_{\A \times \A}$ is surjective, there exist $x_1$ and
    $x_2$ in $\A$ wuth $f_{\A \times \A}(x_1, x_2) = [\A]$.  However, by Lemma
    \ref{card_mult_idemp_l_lt}, we have $f_{\A \times \A}(x_1, x_2) < [\A]$, a
    contradiction.  Thus, $[\A \times \A] \leq [\A]$.
\end{proof}

\begin{lemma} \label{card_idemp}
    $|\A| = |\A \times \A|$.
\end{lemma}
\begin{proof}
    The proof will be by antisymmetry.  The function $f(x) = (x, x)$ is
    injective, showing that $|\A| \leq |\A \times \A|$.  By Lemma
    \ref{card_mult_idemp_A_leq}, we have $\A \leq \A \times \A$, so by Theorem
    \ref{ord_to_card_le} we have $|\A| \leq |\A \times \A|$ as well.
\end{proof}

Now, drop all the assumptions of the previous lemmas and include them as part of
their conditions.

\begin{lemma} \label{ord_mult_idemp}
    For all infinite order types $\A$, we have $|\A| = |\A \times \A|$.
\end{lemma}
\begin{proof}
    By transfinite induction we can get the inductive hypothesis the previous
    lemmas require.  There will be two cases: when a $\beta < [\A]$ with
    $|\beta| = |\A|$ exists, and when no such $\beta$ exists.  In the case where
    $\beta$ exists, the result follows by Lemma \ref{not_card_idemp}, and in the
    case where no such $\beta$ exists, the result follows by Lemma
    \ref{card_idemp}.
\end{proof}

We can finally get to the main result:

\begin{theorem} \label{card_mult_idemp}
    For all infinite cardinals $\kappa$, we have $\kappa\kappa = \kappa$.
\end{theorem}
\begin{proof}
    Let $\A$ be a type with cardinality $\kappa$.  By the well ordering theorem,
    it can be well ordered.  Then the result follows by Lemma
    \ref{ord_mult_idemp}.
\end{proof}

From this result, several other similar results follow.

\begin{theorem} \label{card_plus_idemp}
    For all infinite cardinals $\kappa$, we have $\kappa + \kappa = \kappa$.
\end{theorem}
\begin{proof}
    First, $\kappa + \kappa = 2\kappa$.  Since $\kappa$ is infinite, we have $2
    \leq \kappa$.  Then multiplying by $\kappa$, we get $2\kappa \leq
    \kappa\kappa = \kappa$.  Multiplying $1 \leq 2$ by $\kappa$ we get $\kappa
    \leq 2\kappa$.  Thus, by antisymmetry, $2\kappa = \kappa$.
\end{proof}

\begin{theorem} \label{card_plus_lmax}
    For all cardinals $\kappa$ and $\mu$, if $\kappa$ is infinite and $\mu \leq
    \kappa$, then $\kappa + \mu = \kappa$.
\end{theorem}
\begin{proof}
    From $\mu \leq \kappa$, adding $\kappa$ on both sides we get $\kappa + \mu
    \leq \kappa + \kappa = \kappa$.  From $0 \leq \mu$ and adding $\kappa$ on
    both sides we get $\kappa \leq \kappa + \mu$.  Thus, by antisymmetry we get
    $\kappa + \mu = \kappa$.
\end{proof}

\begin{theorem} \label{card_mult_lmax}
    For all cardinals $\kappa$ and $\mu$, if $\kappa$ is infinite, $0 \neq \mu$,
    and $\mu \leq \kappa$, then $\kappa \mu = \kappa$.
\end{theorem}
\begin{proof}
    From $\mu \leq \kappa$, multiplying by $\kappa$ on both sides we get $\kappa
    \mu \leq \kappa \kappa = \kappa$,  And from $0 \neq \mu$ we get $1 \leq
    \mu$, and multiplying by $\kappa$ we get $\kappa \leq \kappa \mu$.  Thus, by
    antisymmetry we get $\kappa \mu = \kappa$.
\end{proof}

\begin{theorem} \label{card_plus_max}
    For all infinite cardinals $\kappa$ and $\mu$, $\kappa + \mu = \max(\kappa,
    \mu)$.
\end{theorem}
\begin{proof}
    The result follows by looking at the cases $\mu \leq \kappa$ and $\kappa
    \leq \mu$ and applying Theorem \ref{card_plus_lmax}.
\end{proof}

\begin{theorem} \label{card_mult_max}
    For all infinite cardinals $\kappa$ and $\mu$, $\kappa \mu = \max(\kappa,
    \mu)$.
\end{theorem}
\begin{proof}
    The result follows by looking at the cases $\mu \leq \kappa$ and $\kappa
    \leq \mu$ and applying Theorem \ref{card_mult_lmax}.
\end{proof}

\begin{theorem} \label{card_inf_plus_mult}
    For all infinite cardinals $\kappa$ and $\mu$, $\kappa + \mu = \kappa\mu$.
\end{theorem}
\begin{proof}
    By the previous two theorems, both sides are equal to $\max(\kappa, \mu)$.
\end{proof}

With these theorems, we can now easily prove that cardinal operations with
natural numbers are cancellative.

\begin{theorem} \label{card_nat_plus_lcancel} \label{card_nat_plus_rcancel}
    For all cardinals $\mu$ and $\kappa$ and natural numbers $n$ with $n + \mu =
    n + \kappa$, we have $\mu = \kappa$.
\end{theorem}
\begin{proof}
    There are three cases: when both $\mu$ and $\kappa$ are infinite, when one
    is infinite and the other finite, and when both are finite.
    \setcounter{case}{0}
    \begin{case} \textit{$\mu$ and $\kappa$ are both infinite.}
        In this case, we have $n + \mu = \mu$ and $n + \kappa = \kappa$, so $\mu
        = \kappa$.
    \end{case}
    \begin{case} \textit{One of $\mu$ and $\kappa$ is infinite and the other is
        finite.}
        Without loss of generality, assume that $\mu$ is infinite and $\kappa$
        is finite.  Then $n + \mu = \mu = n + \kappa$.  But $\mu$ is infinite
        and $n + \kappa$ is finite, so this case is impossible.
    \end{case}
    \begin{case} \textit{$\mu$ and $\kappa$ are both finite.}
        The result follows from the corresponding property in the natural
        numbers.
    \end{case}
\end{proof}

\begin{theorem} \label{card_nat_mult_lcancel} \label{card_nat_mult_rcancel}
    For all cardinals $\mu$ and $\kappa$ and natural numbers $n$ with $n \mu =
    n \kappa$, we have $\mu = \kappa$.
\end{theorem}
\begin{proof}
    The proof is the same in form as the previous proof.
\end{proof}

\section{Cardinals and Sets}

Recall that sets are thought of as predicates $\U \to \Prop$ on types.
Cardinals interact with them in many ways.  While this was already done above,
whenever $S$ is a set, let $|S|$ denote the cardinality of $\T(S)$.

\begin{theorem} \label{card_le_sub}
    For all cardinals $\kappa$ and types $\A$ with $\kappa \leq |\A|$, there
    exists a set $S : \A \to \Prop$ with $|S| = \kappa$.
\end{theorem}
\begin{proof}
    Let $\U$ be a type with cardinality $\kappa$.  Because $|\U| \leq |\A|$, we
    have an injective function $f : \U \to \A$.  Let $S$ be the image of $f$.
    $f$ is then a bijective function from $\U$ to $S$, showing that $|S| =
    |\U|$.
\end{proof}

\begin{theorem} \label{card_subs_le}
    For all sets $A$ and $B$ on a type $\U$, if $A \subseteq B$, then $|A| \leq
    |B|$.
\end{theorem}
\begin{proof}
    The identity function is an injective function from $A$ to $B$.
\end{proof}

\begin{theorem} \label{card_all}
    For all types $\U$, $|\U| = |\bm U|$, where $\bm U$ is the universal set.
\end{theorem}
\begin{proof}
    The identity function is bijective.
\end{proof}

\begin{theorem} \label{card_sub_le}
    For all types $\A$ and sets $S : \A \to \Prop$, we have $|S| \leq |\A|$.
\end{theorem}
\begin{proof}
    Since $S \subset \bm U$, we have $|S| \leq |\bm U| = |\A|$.
\end{proof}

\begin{theorem} \label{card_minus_le}
    For all sets $A$ and $B$, $|A - B| \leq |A|$.
\end{theorem}
\begin{proof}
    Direct from Theorem \ref{card_subs_le}.
\end{proof}

\begin{theorem} \label{image_le}
    For all types $\U$ and $\V$ and functions $f : \U \to \V$, with $S$ the
    image of $f$, we have $|S| \leq |\U|$.
\end{theorem}
\begin{proof}
    By the partition principle, it suffices to find a surjective function from
    the image of $\U$ to the image of $f$.  $f$ is precisely such a function.
\end{proof}

\begin{theorem} \label{image_under_le}
    For all types $\U$ and $\V$, sets $A : \U \to \Prop$, and functions $f : \U
    \to \V$, we have $|f(A)| \leq |A|$.
\end{theorem}
\begin{proof}
    The proof is the the same in form as the previous proof.
\end{proof}

\begin{theorem} \label{card_union_left}
    For all sets $S$ and $T$, $|S| \leq |S \cup T|$.
\end{theorem}
\begin{proof}
    Direct from Theorem \ref{card_subs_le}.
\end{proof}

\begin{theorem} \label{card_union_right}
    For all sets $S$ and $T$, $|T| \leq |S \cup T|$.
\end{theorem}
\begin{proof}
    Direct from Theorem \ref{card_subs_le}.
\end{proof}

\begin{theorem} \label{card_plus_union}
    For all sets $S$ and $T$, $|S \cup T| \leq |S| + |T|$.
\end{theorem}
\begin{proof}
    Define a function $f : (S \cup T) \to (S + T)$ given by
    \[
        f(x) = \begin{cases}
            x| & \text{if $x \in S$} \\
            |x & \text{if $x \in T$.}
        \end{cases}
    \]
    Let $a$ and $b$ be such that $f(a) = f(b)$.  The only possible cases are $a|
    = b|$ and $|a = |b$, and in both cases we have $a = b$.
\end{proof}

\begin{theorem} \label{zero_is_empty}
    For all sets $S$, if $0 = |S|$, then $S = \varnothing$.
\end{theorem}
\begin{proof}
    Direct from Theorem \ref{empty_eq}.
\end{proof}

\begin{theorem} \label{empty_set_size}
    $|\varnothing| = 0$.
\end{theorem}
\begin{proof}
    Direct from Theorem \ref{empty_set_size}.
\end{proof}

\begin{theorem} \label{singleton_set_size}
    For all types $\U$ and $a : \U$, $|\{a\}| = 1$.
\end{theorem}
\begin{proof}
    Defining $f : \{a\} \to \S$ as $f(a) = I$ and $g : \S \to \{a\}$ as $g(I) =
    a$, $f$ and $g$ are inverses of each other.
\end{proof}

\begin{theorem} \label{all_greater_inf_set}
    For ordered sets $\U$ and sets $S : \U \to \Prop$, if $S$ is nonempty and if
    for all $a \in S$, there exists a $b \in S$ with $a < b$, then $|S|$ is
    infinite.
\end{theorem}
\begin{proof}
    This is just a special case of Theorem \ref{all_greater_inf}.
\end{proof}

\begin{theorem} \label{card_plus_one}
    For all sets $S$ and $a \in S$, $|S| = |S - \{a\}| + 1$.
\end{theorem}
\begin{proof}
    Define functions $f : S \to (S - \{a\}) + \S$ and $g : (S - \{a\}) + \S \to
    S$ given by
    \[
        f(x) = \begin{cases}
            |I & \text{if $x = a$} \\
            x| & \text{if $x \neq a$}
        \end{cases}
    \]
    and
    \[
        g(x) = \begin{cases}
            y & \text{if $x = y|$} \\
            a & \text{if $x = |I$.}
        \end{cases}
    \]
    Then when $x = y|$,
    \[
        f(g(x)) = f(y) = y| = x,
    \]
    and when $x = |I$,
    \[
        f(g(x)) = f(a) = |I = x,
    \]
    so $f(g(x)) = x$ for all $x$.  And when $x = a$,
    \[
        g(f(x)) = g(|I) = a = x,
    \]
    and when $x \neq a$,
    \[
        g(f(x)) = g(x|) = x,
    \]
    so $g(f(x)) = x$ for all $x$.
\end{proof}

\begin{theorem} \label{card_plus_one_nat}
    For all sets $S$, natural numbers $n$, and $a \in S$, if $|S| = n + 1$, then
    $|S - \{a\}| = n$.
\end{theorem}
\begin{proof}
    By the previous theorem, we have $|S| = |S - \{a\}| + 1$, so $|S - \{a\}| +
    1 = n + 1$.  By Theorem \ref{card_nat_plus_lcancel}, we can cancel the $1$
    to get $|S - \{a\}| = n$.
\end{proof}

\begin{theorem} \label{inter_le}
    For all collections $\mathcal S : (\U \to \Prop) \to \Prop$, sets $S \in
    \mathcal S$, and cardinals $\mu$, if $|S| \leq \mu$, then $|\bigcap \mathcal
    S| \leq \mu$.
\end{theorem}
\begin{proof}
    \begin{align*}
        \bigcap \mathcal S &\subseteq S \\
        \left|\bigcap \mathcal S\right| &\leq |S| \leq \mu.
    \end{align*}
\end{proof}

\begin{theorem} \label{inter_lt}
    For all collections $\mathcal S : (\U \to \Prop) \to \Prop$, sets $S \in
    \mathcal S$, and cardinals $\mu$, if $|S| < \mu$, then $|\bigcap \mathcal
    S| < \mu$.
\end{theorem}
\begin{proof}
    \begin{align*}
        \bigcap \mathcal S &\subseteq S \\
        \left|\bigcap \mathcal S\right| &\leq |S| < \mu.
    \end{align*}
\end{proof}

\begin{theorem} \label{finite_inter_finite}
    For all collections $\mathcal S : (\U \to \Prop) \to \Prop$, if there exists
    an $S \in \mathcal S$ that is finite, then $\bigcap \mathcal S$ is finite as
    well.
\end{theorem}
\begin{proof}
    Special case of Theorem \ref{inter_lt}.
\end{proof}

\begin{theorem} \label{union_size_mult}
    For all collections $\mathcal S : (\U \to \Prop) \to \Prop$ and cardinals
    $\kappa$ and $\mu$, if $|\mathcal S| \leq \kappa$ and $|S| \leq \mu$ for all
    $S \in \mathcal S$, then $|\bigcup \mathcal S| \leq \kappa\mu$.
\end{theorem}
\begin{proof}
    Let $\A$ and $\B$ be types with cardinality $\kappa$ and $\mu$ respectively.
    Because $|\mathcal S| \leq |\A|$, we have an injective function $f :
    \mathcal S \to \A$.  Because $|S| \leq |\B|$ for all $S \in \mathcal S$, we
    have injective functions $g_S : S \to \B$ for all $S \in \mathcal S$.
    Define a function $G : \bigcup \mathcal S \to \mathcal S$ that takes an $x
    \in \mathcal S$ and produces a set $S \in \mathcal S$ that contains $x$.
    Then define a function $h : \bigcup \mathcal S \to \A \times \B$ given by
    $h(x) = (f(G(x)), g_{G(x)}(x))$.  We will prove that $h$ is injective.

    Let $a$ and $b$ be such that $h(a) = h(b)$.  Then $f(G(a)) = f(G(b))$ and
    $g_{G(a)}(a) = g_{G(b)}(b)$.  Because $f$ is injective, we have $G(a) =
    G(b)$.  Thus, $g_{G(a)}(a) = g_{G(a)}(b)$, and because $g_{G(a)}$ is
    injective, we have $a = b$ as required.
\end{proof}

As an immediate application of this theorem, we have this important result:

\begin{theorem} \label{countable_union_countable}
    The countable union of countable sets is countable.  More explicitly, for
    all collections $\mathcal S$ with $|\mathcal S| \leq |\N|$, if $|S| \leq
    |\N|$ for all $S \in \mathcal S$, then $|\bigcup \mathcal S| \leq |\N|$.
\end{theorem}
\begin{proof}
    By Theorem \ref{union_size_mult}, we have $|\bigcup \mathcal S| \leq
    |\N|\,|\N|$, and then the result follows by Theorem \ref{card_mult_idemp}.
\end{proof}

Theorem \ref{union_size_mult} used non-strict inequalities.  Does the result
hold for strict inequalities too?

\begin{theorem}
    It is not the case that for all collections $\mathcal S : (\U \to \Prop) \to
    \Prop$ and cardinals $\kappa$ and $\mu$, if $|\mathcal S| < \kappa$ and $|S|
    < \mu$ for all $S \in \mathcal S$, then $|\bigcup \mathcal S| < \kappa\mu$.
\end{theorem}
\begin{proof}
    Let $\U$ be some type with cardinality $\aleph_\omega$.  By Theorem
    \ref{card_le_sub}, $\U$ has subsets of all cardinalities $\aleph_0$,
    $\aleph_1$, etc.  Let $b : \N \to (\U \to \Prop)$ be a function that
    produces such a subset for each cardinality, that is, $|b(0)| = \aleph_0$,
    $|b(1)| = \aleph_1$, etc.  Then let $\mathcal S$ be the image of $b$.  By
    Theorem \ref{image_le}, $|\mathcal S| \leq |\N| = \aleph_0 < \aleph_\omega$,
    and by construction, for all $S \in \mathcal S$, we have $|S| <
    \aleph_\omega$.  Now, $|\bigcup \mathcal S| > \aleph_n$ for all natural
    numbers $n$ by construction, so by Theorem \ref{aleph_least} we have
    $|\bigcup \mathcal S| \geq \aleph_\omega$.  Since $\aleph_\omega
    \aleph_\omega = \aleph_\omega$ by Theorem \ref{card_mult_idemp}, we have our
    counterexample.
\end{proof}

While the strict version of the theorem is not universally true, it is true in
many cases, including the following important case:

\begin{theorem} \label{finite_union_finite}
    The finite union of finite sets is finite, that is, for all collections
    $\mathcal S : (\U \to \Prop) \to \Prop$, if $|\mathcal S| < |\N|$ and $|S| <
    |\N|$ for all $S \in \mathcal S$, then $|\bigcup \mathcal S| < |\N|$.
\end{theorem}
\begin{proof}
    First, if $\mathcal S$ is empty, we have $|\bigcup \varnothing| =
    |\varnothing| = 0 < |\N|$, so from now on assume that $\mathcal S$ is
    nonempty.  Let $g : \mathcal S \to \N$ be the function that gets the size of
    each set in $\mathcal S$.  Because $\mathcal S$ is finite, the image of $g$
    is finite too by Theorem \ref{image_le}, so by Theorem \ref{finite_max} $g$
    takes on a maximum value $n$.  Thus, for all sets $S \in \mathcal S$, we
    have $|S| \leq n$.  Furthermore, because $\mathcal S$ is finite, we have a
    natural number $m$ such that $|\mathcal S| \leq m$.  Thus, by Theorem
    \ref{union_size_mult}, we have $|\bigcup \mathcal S| \leq mn < |\N|$,
    showing that $\bigcup \mathcal S$ is finite.
\end{proof}

\section{The Cardinality of Previously Developed Types}

Throughout this chapter, the only concrete type we have found the cardinality of
is the natural numbers.  In this section, we will look at the cardinality of
several of the other types we have looked at already.

\begin{theorem} \label{equiv_card_le}
    If $\U$ is a type with an equivalence relation $\sim$, then $|\U/{\sim}|
    \leq |\U|$.
\end{theorem}
\begin{proof}
    The function $\{X\}$ given by Definition \ref{from_equiv} is a function from
    $\U/{\sim}$ to $\U$.  If we have $\{a\} = \{b\}$, then $[\{a\}] = [\{b\}]$,
    and by Theorem \ref{from_equiv_eq} we have $a = b$, showing that the
    function is injective.
\end{proof}

\begin{theorem} \label{int_size}
    $|\Z| = |\N|$.
\end{theorem}
\begin{proof}
    Since the integers are defined as equivalence classes of pairs of natural
    numbers, we have $|\Z| \leq |\N| \, |\N| = |\N|$, and $\iN$ is injective, so
    $|\N| \leq |\Z|$.  Thus, by antisymmetry, $|\Z| = |\N|$.
\end{proof}

\begin{theorem} \label{rat_size}
    $|\Q| = |\N|$.
\end{theorem}
\begin{proof}
    Since the rational numbers are defined as equivalence classes of pairs of an
    integer and a nonzero integer, we have
    \[
        |\Q| \leq |\Z| \, |\Z - \{0\}| \leq |\Z| \, |\Z| = |\Z| = |\N|,
    \]
    and $\iN$ is injective, so $|\N| \leq |\Q|$.  Thus, by antisymmetry, $|\Q| =
    |\N|$.
\end{proof}

\begin{theorem} \label{dense_open_infinite}
    In any dense ordered set $\U$, if $a < b$, then the open interval $(a, b)$
    is infinite.
\end{theorem}
\begin{proof}
    The proof will be by Theorem \ref{all_greater_inf_set}.  $(a, b)$ is
    nonempty because $\U$ is dense.  Now if $x \in (a, b)$, by $\U$ being dense
    we have a $y$ with $x < y < b$.  Then $x < y$ and $y \in (a, b)$, so $(a,
    b)$ is infinite by Theorem \ref{all_greater_inf_set}.
\end{proof}

\begin{theorem} \label{dense_closed_infinite}
    In any dense ordered set $\U$, if $a < b$, then the closed interval $[a, b]$
    is infinite.
\end{theorem}
\begin{proof}
    Since $(a, b) \subseteq [a, b]$, the theorem follows by a direct application
    of Theorem \ref{card_subs_le}.
\end{proof}

\begin{theorem} \label{arch_ordered_size}
    The cardinality of an Archimedean ordered field is at most $2^{|\N|}$.
\end{theorem}
\begin{proof}
    Let $\U$ be an Archimedean ordered field.  We have $2^{|\N|} = 2^{|\Q|} =
    |\Q \to \Prop|$, so it suffices to find an injective function from $\U$ to
    $\Q \to \Prop$.  Define $f : \U \to (\Q \to \Prop)$ by $f(x) = \{q : \Q \mid
    q < x\}$.  Now let $f(a) = f(b)$.  If $a < b$, by Theorem
    \ref{rat_dense_in_arch}, we have a rational number $r$ with $a < r < b$.
    This means that $r \in f(b)$ but $r \notin f(a)$, contradicting $f(a) =
    f(b)$.  Thus, we can't have $a < b$, and by the same argument, we can't have
    $b < a$.  This means that $a = b$, so $f$ is injective.
\end{proof}

\begin{theorem} \label{empty_list_card}
    For all types $\U$, if $\U$ has no values, then $|\L(\U)| = 1$.
\end{theorem}
\begin{proof}
    The only possible list is $[]$, so the function defined by $f([]) = I$ is a
    bijection between $\L(\U)$ and $\S$.
\end{proof}

\begin{theorem} \label{list_infinite}
    For all nonempty types $\U$, $|\L(\U)| \geq |\N|$.
\end{theorem}
\begin{proof}
    Let $x$ be a value in $\U$.  Then define a function $f$ from $\N$ to
    $\L(\U)$ defined by $f(n) = [x]_n$.  This is an injection.
\end{proof}

\begin{theorem} \label{list_card_le}
    For all types $\U$, we have $|\U| \leq |\L(\U)|$.
\end{theorem}
\begin{proof}
    The function bringing $x : \U$ to $[x]$ is an injection.
\end{proof}

\begin{theorem} \label{list_size_card}
    Let $\U$ be a type and $n$ a natural number.  Denote the set of lists of
    size $n$ as $\L_n(\U)$.  Then $|\L_n(\U)| = |\U|^n$.
\end{theorem}
\begin{proof}
    The proof will be by induction on $n$.  When $n = 0$, we must prove that
    $|\L_0(\U)| = |\U|^0 = 1$.  The only list of length $0$ is the empty list, so
    that set has cardinality $1$.

    Now assume that $|\L_n(\U)| = |\U|^n$.  Then $|\U|^{S(n)} = |\U|^n |\U| =
    |\L_n(\U)| \, |\U|$, so we must find a bijection between $\L_{S(n)}(\U)$ and
    $\L_n(\U) \times \U$.  Given a list $(x : a) : \L_{S(n)}(\U)$, we can bring
    it to the pair $(a, x)$.  This mapping is a bijection.
\end{proof}

\begin{theorem} \label{finite_list_card}
    For all finite nonempty types $\U$, we have $|\L(\U)| = |\N|$.
\end{theorem}
\begin{proof}
    Define a collection $\mathcal S : (\L(\U) \to \Prop) \to \Prop$ given by
    \[
        \mathcal S = \{\L_0(\U), \L_1(\U), \L_2(\U), \ldots\}.
    \]
    The cardinality of $\bigcup \mathcal S$ is equal to the cardinality of
    $|\L(\U)|$, so it suffices to prove that $|\bigcup \mathcal S| = |\N|$.  We
    already have $|\N| \leq |\L(\U)|$ by Theorem \ref{list_infinite}, so by
    antisymmetry it suffices to prove that $|\bigcup \mathcal S| \leq |\N|$.
    The set $\mathcal S$ itself is countable infinite by construction.  Given
    any $\L_n(\U) \in \mathcal S$, we have $|\L_n(\U)| = |\U|^n$ by Theorem
    \ref{list_size_card}, and since $|\U|$ and $n$ are both finite, $|\U|^n$ is
    finite as well.  Thus, $\mathcal S$ is a countable union of finite sets, so
    by Theorem \ref{countable_union_countable}, $|\bigcup \mathcal S|$ is
    countable as well.
\end{proof}

\begin{theorem} \label{infinite_list_card}
    For all infinite types $\U$, we have $|\L(\U)| = |\U|$.
\end{theorem}
\begin{proof}
    Most of the details of this proof are the same as the previous proof so the
    details will be ommitted.  We already have $|\U| \leq |\L(\U)|$ by Theorem
    \ref{list_card_le}.  Furthermore, because $\U$ is infinite, we have $|\U|^n
    = |\U|$ for all natural numbers $n$.  Then with the same argument as in the
    previous proof, by Theorem \ref{union_size_mult} we have $|\L(\U)| \leq |\N|
    \, |\U| = |\U|$.
\end{proof}

\section{Large Countable Ordinals}

This topic is more about ordinals, but it requires some stuff about the
cardinals, so I'm putting this here.  The ordinal chapter got to have stuff
about the cardinals anyway, so this is just payback.

Large countable ordinals is a massive topic, and I haven't even come close to
scratching the surface of how big we can get.  I only managed to get up to the
Feferman-Sch\"utte ordinal, which is what I will be describing here.  First,
let's look at a bit of the general theory of countable ordinals.  We saw earlier
that we can convert ordinals to cardinals.  We say that an ordinal is countable
if its corresponding cardinal is countable, that is, when $|\alpha| \leq |\N|$.

\begin{theorem} \label{ord_countable_suc}
    For all countable ordinals $\alpha$, $S(\alpha)$ is countable.
\end{theorem}
\begin{proof}
    If $\alpha$ is finite, then $S(\alpha)$ is finite too.  If $|\alpha| =
    |\N|$, then because $|\N| + 1 = |\N|$, $|S(\alpha)| = |\alpha + \S| =
    |\alpha| + 1 = |\N|$.
\end{proof}

\begin{theorem} \label{ord_sup_countable}
    The supremum of a countable set of countable ordinals is countable.
\end{theorem}
\begin{proof}
    Let $S$ be a countable set of countable ordinals.  Note that because we
    can't directly form a cardinal from a set of ordinals, when talking about
    the cardinality of $S$, we are talking about it in terms of Definition
    \ref{small_set_to_card}, which makes this proof much more complicated than
    how it is usually presented.

    First, when $\sup S \in S$, we directly have $|\sup S| \leq |\N|$ because
    $S$ contains countable ordinals, so now consider the case where $\sup S
    \notin S$.  This means that $\alpha < \sup S$ for all $\alpha \in S$.  Let
    $\A$ be a well-ordered type with order type $\sup S$.  We must prove that
    $|\A|$ is countable.

    Consider the collection $\mathcal S : (\A \to \Prop) \to \Prop$ given by
    \[
        \mathcal S = \{\A_{\mathcal O_\A^{-1}(\alpha)} \mid \alpha \in S\},
    \]
    that is, the initial segments in $\A$ of ordinals $\alpha \in S$, with the
    correspondence $\mathcal O_\A$ given by Definition \ref{ord_type_init_ord}.
    Then for any $x : \A$, we know that $\mathcal O_\A(x) < [\A] = \sup S$, so
    by Theorem \ref{ord_sup_in} we have a $\beta \in S$ such that $\mathcal
    O_\A(x) < \beta$.  This means that $x < \mathcal O_\A^{-1}(\beta)$, showing
    that $x \in \A_{\mathcal O_\A^{-1}(\beta)}$, and since $\beta \in S$, we
    have $x \in \bigcup \mathcal S$.  This argument works for any $x : \A$,
    showing that $\bigcup \mathcal S$ is equal to the universal set on $\A$.
    Thus, $|\A| = |\bigcup \mathcal S|$, so we can now apply Theorem
    \ref{countable_union_countable} on $\mathcal S$ to prove $|\A| = |\N|$.

    To prove that $\mathcal S$ is countable, we know that because $S$ is
    countable, we have a type $\X$ that is in bijection with $S$ with $|\X| \leq
    |\N|$.  Let $f : \X \to S$ be such a bijection.  Given a set $\A_{\mathcal
    O_\A^{-1}(\alpha)} \in \mathcal S$, we can produce a value $f^{-1}(\alpha) :
    \X$.  If this mapping is injective, we will have proved that $\mathcal S$ is
    countable.  So let $f^{-1}(\alpha) = f^{-1}(\beta)$, and we must prove that
    $\A_{\mathcal O_\A^{-1}(\alpha)} = \A_{\mathcal O_\A^{-1}(\beta)}$.  By
    applying $f$ to $f^{-1}(\alpha) = f^{-1}(\beta)$, we get $\alpha = \beta$,
    from which the result follows.  Thus, $\mathcal S$ is countable.

    To prove that every $T \in \mathcal S$ is countable, let $\beta \in S$.  We
    must prove that $\A_{\mathcal O_\A^{-1}(\beta)}$ is countable.  Let $\B$ be
    a well-ordered type with order type $\beta$.  By assumption, $|\B|$ is
    countable, so it suffices to prove that $|\A_{\mathcal O_\A^{-1}([\B])}| \leq
    |\B|$.  Define $f : \A_{\mathcal O_\A^{-1}([\B])} \to \B$ by
    \[
        f(a) = \mathcal O_\B^{-1}(\mathcal O_\A(a)).
    \]
    We will prove that $f$ is injective.  Let $f(a) = f(b)$.  Then
    \begin{align*}
        f(a) &= f(b) \\
        \mathcal O_\B^{-1}(\mathcal O_\A(a))
            &= \mathcal O_\B^{-1}(\mathcal O_\A(b)) \\
        \mathcal O_\A(a) &= \mathcal O_\A(b) \\
        a &= b.
    \end{align*}
    Because $f$ is injective, every set in $\mathcal S$ is countable.
\end{proof}

\begin{theorem} \label{ord_countable_lt}
    For all ordinals $\alpha$ and $\beta$ with $\alpha < \beta$, if $\beta$ is
    countable, then $\alpha$ is countable.
\end{theorem}
\begin{proof}
    Trivial.
\end{proof}

\begin{theorem} \label{nat_ord_countable}
    All natural numbers, as ordinals, are countable.
\end{theorem}
\begin{proof}
    Direct consequence of Theorem \ref{ord_to_card_nat}.
\end{proof}

\begin{theorem} \label{omega_countable}
    $\omega$ is countable.
\end{theorem}
\begin{proof}
    $|\omega| = |\N|$.
\end{proof}

\begin{theorem} \label{ord_normal_countable}
    For all ordinals functions $f$ that satisfy the normal limit condition, if
    $f(0)$ is countable and if $f(\alpha)$ being countable implies that
    $f(S(\alpha))$ is countable, then $f(\alpha)$ is countable for all countable
    $\alpha$.
\end{theorem}
\begin{proof}
    The proof will be by ordinal induction.  The zero and successor cases follow
    almost directly by assumption.  The limit case follows from Theorem
    \ref{ord_sup_countable}.
\end{proof}

\begin{theorem} \label{ord_countable_plus}
    If $\alpha$ and $\beta$ are countable, then $\alpha + \beta$ is countable.
\end{theorem}
\begin{proof}
    Because adding $\alpha$ on the left satisfies the normal limit condition, we
    can use Theorem \ref{ord_normal_countable}.  When $\beta = 0$, we have
    $\alpha + 0 = \alpha$, which is countable by assumption.  When $\alpha +
    \beta$ is countable, we have $S(\alpha + \beta) = \alpha + S(\beta)$ being
    countable by Theorem \ref{ord_countable_suc}.
\end{proof}

\begin{theorem} \label{ord_countable_mult}
    If $\alpha$ and $\beta$ are countable, then $\alpha + \beta$ is countable.
\end{theorem}
\begin{proof}
    Because multiplying by $\alpha$ on the left satisfies the normal limit
    condition, we can use Theorem \ref{ord_normal_countable}.  When $\beta = 0$,
    we have $\alpha 0 = 0$, which is countable.  When $\alpha \beta$ is
    countable, we have $\alpha \beta + \alpha = \alpha S(\beta)$ being countable
    by Theorem \ref{ord_countable_plus}.
\end{proof}

\begin{theorem} \label{ord_countable_pow}
    If $\alpha$ and $\beta$ are countable, then $\alpha ^ \beta$ is countable.
\end{theorem}
\begin{proof}
    We will need a special case when $\alpha = 0$.  But here, we have either
    $0^\beta = 0$ or $0^\beta = 1$, and these are both countable.  When $\alpha
    \neq 0$, because ordinal exponentiation with $\alpha \neq 0$ satisfies the
    normal limit condition, we can use Theorem \ref{ord_normal_countable}.  When
    $\beta = 0$, we have $\alpha^0 = 1$, which is countable.  When
    $\alpha^\beta$ is countable, we have $\alpha^\beta \alpha =
    \alpha^{S(\beta)}$ being countable by Theorem \ref{ord_countable_mult}.
\end{proof}

\begin{definition}
    Define the ordinal $\varepsilon_0$ as the supremum of the set
    \[
        \{1, \omega, \omega^\omega, \omega^{\omega^\omega}, \ldots\}.
    \]
\end{definition}

\begin{lemma} \label{omega_epsilon0}
    $\omega^{\varepsilon_0} = \varepsilon_0$.
\end{lemma}
\begin{proof}
    Let ${}^n\omega$ be a power tower of $n$ $\omega$s.  Then
    \[
        \varepsilon_0 = \sup\{{}^n\omega \mid n < \omega\}.
    \]
    This means that
    \begin{align*}
        \omega^{\varepsilon_0}
        &= \omega^{\sup\{{}^n\omega \mid n < \omega\}} \\
        &= \sup\{{}^{S(n)}\omega \mid n < \omega\}.
    \end{align*}
    So we must prove that
    \[
        \sup\{{}^{S(n)}\omega \mid n < \omega\},
        = \sup\{{}^n\omega \mid n < \omega\}
    \]
    which we will do by antisymmetry and Theorem \ref{ord_sup_leq_sup}.  The
    reverse direction is trivial.  For the forward direction, let $m < \omega$.
    Then setting $n = S(m)$ in the second set, we get \[\sup\{{}^{S(m)}\omega
    \mid n < \omega\} \leq \sup\{{}^{S(m)}\omega \mid n < \omega\}\] as
    required.
\end{proof}

\begin{theorem} \label{epsilon0_countable}
    $\varepsilon_0$ is countable.
\end{theorem}
\begin{proof}
    By induction and Theorem \ref{ord_countable_pow}, ${}^n\omega$ is countable
    for all $n < \omega$.  Thus by Theorem \ref{ord_sup_countable},
    $\varepsilon_0 = \sup\{{}^n\omega \mid n < \omega\}$ is countable.
\end{proof}

\begin{theorem} \label{epsilon0_least}
    For all $\alpha$ with $\omega^\alpha = \alpha$, we have $\varepsilon_0 \leq
    \alpha$.
\end{theorem}
\begin{proof}
    It suffices to prove that $\alpha$ is an upper bound of $\{{}^n\omega\}$.
    So let $n$ be a natural number.  We will prove that ${}^n\omega \leq \alpha$
    by induction on $n$.  When $n = 0$, this reduces to $1 \leq \alpha$, which
    is equilavent to $0 < \alpha$.  If $0 = \alpha$, we would have $\omega^0 = 1
    = 0$, which is impossible, so we must have $0 \neq \alpha$, hence $0 <
    \alpha$.  So the base case is true.

    Now assume that ${}^n\omega \leq \alpha$.  Then by Theorem \ref{ord_pow_le},
    \begin{align*}
        \omega^{{}^n\omega} &\leq \omega^\alpha \\
        {}^{S(n)}\omega &\leq \alpha.
    \end{align*}
\end{proof}

\begin{theorem} \label{ord_normal_family_fixed_set_card}
    For all small families $f_i$ with index type $\X$, we have $|F(\alpha)| \leq
    |\L(\X)|$.
\end{theorem}
\begin{proof}
    We have a bijection $g : \X \to f_i$.  Given a list of functions $[f_0, f_1,
    \ldots, f_n]$, we can take them to a list in $\X$ $[g^{-1}(f_0),
    g^{-1}(f_1), \ldots, g^{-1}(f_n)]$.  This process is injective by $g$ being
    bijective, showing that $|F(\alpha)| \leq |\L(\X)|$.
\end{proof}

\begin{lemma} \label{small_set_list_countable}
    For all types $\X$ and small sets $S$ on $\X$, if $S$ is countable, then
    with $\Y$ a type in bijection with $S$, we have $|\L(\Y)| \leq |\N|$.
\end{lemma}
\begin{proof}
    Because $S$ is countable, $|\Y|$ is countable.  When $\Y$ is empty, the
    result follows from Theorem \ref{empty_list_card}.  When $\Y$ is nonempty
    and finite, the result follows from Theorem \ref{finite_list_card}.  When
    $\Y$ is (countably) infinite, the result follows from Theorem
    \ref{infinite_list_card}.
\end{proof}

\begin{theorem} \label{ord_normal_family_fixed_countable}
    Let $f_i$ be a countable family of normal functions.  If $f_x(\beta)$ is
    countable for all countable $\beta$ and all $x$, then for any countable
    $\alpha$, then $\sup F(\alpha)$, the first fixed point of $f_i$ greater than
    or equal to $\alpha$, is countable.
\end{theorem}
\begin{proof}
    By Lemma \ref{small_set_list_countable}, $F(\alpha)$ is countable.  Then by
    Theorem \ref{ord_sup_countable}, it suffices to prove that every $\delta \in
    F(\alpha)$ is countable.  Every such $\delta$ is equal to various $f_i$
    functions applied to $\alpha$.  Because applying each of these functions to
    a countable ordinal produces a countable ordinal, and because $\alpha$ is
    countable, inductively, $\delta$ is countable as well.
\end{proof}

\begin{theorem} \label{ord_family_derivative_countable}
                \label{ord_derivative_countable}
    Let $f_i$ be a countable family of normal functions.  If $f_x(\beta)$ is
    countable for all countable $\beta$ and all $x$, then for any countable
    $\alpha$, $f'(\alpha)$ is countable.
\end{theorem}
\begin{proof}
    Because $f'$ is normal, we can use Theorem \ref{ord_normal_countable}.
    First, $f'(0)$ is equal to $\sup F(0)$, which is countable by Theorem
    \ref{ord_normal_family_fixed_countable}.  Then when $f'(\alpha)$ is
    countable, $f'(S(\alpha)) = \sup F(S(f'(\alpha)))$, which is countable by
    Theorems \ref{ord_normal_family_fixed_countable} and
    \ref{ord_countable_suc}.
\end{proof}

\begin{definition}
    Let $\alpha$ be an ordinal.  We will define the veblen function
    $\varphi_\alpha : \Ord \to \Ord$ by ordinal recursion such that
    \begin{align*}
        \varphi_0(\beta) &= \omega^\beta \\
        \varphi_{S(\alpha)}(\beta) &= \varphi'_\alpha(\beta) \\
        \varphi_\gamma(\beta) &= \varphi'_{\delta < \gamma}(\beta).
    \end{align*}
    In the limit case, the derivative is the derivative of the family of
    functions $\varphi_\delta$ for all $\delta < \gamma$.
\end{definition}

By definition $\varphi_\alpha$ is a normal function for all $\alpha$.

\begin{theorem}
    $\varphi_1(0) = \varepsilon_0$.
\end{theorem}
\begin{proof}
    Both $\varphi_1(0)$ and $\varepsilon_0$ are the least fixed point of
    $\omega^\beta$ by Theorems \ref{ord_normal_family_fixed_least} and
    \ref{epsilon0_least}, so they are equal.
\end{proof}

\begin{lemma} \label{veblen_gt_eq_nat}
    For all ordinals $\alpha$ and $\zeta$ and nonzero natural numbers $n$, we
    have
    \[
        \varphi_\alpha(\varphi_{\alpha + n}(\zeta))
        = \varphi_{\alpha + n}(\zeta).
    \]
\end{lemma}
\begin{proof}
    The proof will be by induction on $n$.  Because $n$ is nonzero, the base
    case is $n = 1$.  Then the result follows by $\varphi$ being defined on
    successors as the derivative of the previous $\varphi$ function.  Now assume
    that for all $\alpha$ and $\zeta$,
    \[
        \varphi_\alpha(\varphi_{\alpha + n}(\zeta))
        = \varphi_{\alpha + n}(\zeta).
    \]
    Then
    \begin{align*}
        \varphi_\alpha(\varphi_{\alpha + S(n)}(\zeta))
        &= \varphi_\alpha(\varphi_{S(\alpha + n)}(\zeta)) \\
        &= \varphi_\alpha(\varphi_{\alpha + n}(\varphi_{S(\alpha + n)}(\zeta)))\\
        &= \varphi_{\alpha + n}(\varphi_{S(\alpha + n)}(\zeta)) \\
        &= \varphi_{S(\alpha + n)}(\zeta) \\
        &= \varphi_{\alpha + S(n)}(\zeta).
    \end{align*}
\end{proof}

\begin{lemma} \label{veblen_gt_eq_lim}
    For all ordinals $\alpha$ and $\beta$ with $\alpha < \beta$, if $\beta$ is a
    limit ordinal, then
    \[
        \varphi_\alpha(\varphi_\beta(\zeta)) = \varphi_\beta(\zeta)
    \]
    for all ordinals $\zeta$.
\end{lemma}
\begin{proof}
    This is true by definition.
\end{proof}

\begin{theorem} \label{veblen_gt_eq}
    For all ordinals $\alpha$ and $\beta$ with $\alpha < \beta$,
    \[
        \varphi_\alpha(\varphi_\beta(\zeta)) = \varphi_\beta(\zeta)
    \]
    for all ordinals $\zeta$.
\end{theorem}
\begin{proof}
    First, if there exists an $n$ with $\alpha + n = \beta$, the result follows
    from Lemma \ref{veblen_gt_eq_nat}.  So now consider the case when no such
    $n$ exists.  By Theorems \ref{ord_near_lim} and \ref{ord_near_lim_lim}, let
    $\gamma$ be the limit ordinal such that $\alpha < \gamma$ and let $n$ be the
    natural number such that $\gamma + n = \beta$.  Now if $n = 0$, the result
    follows directly from Lemma \ref{veblen_gt_eq_lim}.  If $n \neq 0$,
    \begin{align*}
        \varphi_\alpha(\varphi_\beta(\zeta))
        &= \varphi_\alpha(\varphi_{\gamma + n}(\zeta)) \\
        &= \varphi_\alpha(\varphi_\gamma(\varphi_{\gamma + n}(\zeta))) \\
        &= \varphi_\gamma(\varphi_{\gamma + n}(\zeta)) \\
        &= \varphi_{\gamma + n}(\zeta) \\
        &= \varphi_{\beta}(\zeta).
    \end{align*}
\end{proof}

\begin{theorem} \label{veblen_le}
    $\varphi_\alpha(0)$, as a function of $\alpha$, is orderly.
\end{theorem}
\begin{proof}
    Let $\alpha \leq \beta$.  If $\alpha = \beta$, we are done, so let $\alpha <
    \beta$.  Then
    \[
        \varphi_\alpha(0) \leq \varphi_\alpha(\varphi_\beta(0))
        = \varphi_\beta(0).
    \]
\end{proof}

\begin{theorem} \label{veblen_inj}
    $\varphi_\alpha(0)$, as a function of $\alpha$, is injective.
\end{theorem}
\begin{proof}
    Let $\varphi_\alpha(0) = \varphi_\beta(0)$.  For a contradiction, assume
    without loss of generality that $\alpha < \beta$.  Then we would have
    \[
        \varphi_\alpha(0) = \varphi_\beta(0) = \varphi_\alpha(\varphi_\beta(0)),
    \]
    and because $\varphi_\alpha$ is injective, we have $0 = \varphi_\beta(0)$.
    Because $\varphi_\alpha(0)$ is orderly, we have $\varphi_0(0) \leq
    \varphi_\beta(0)$, so $1 \leq 0$, which is impossible.
\end{proof}

\begin{theorem} \label{veblen_normal}
    $\varphi_\alpha(0)$, as a function of $\alpha$, satisfies the normal limit
    condition.
\end{theorem}
\begin{proof}
    Let $\alpha$ be a limit ordinal.  Let $\Phi(\alpha)$ be the set given by
    Definition \ref{ord_normal_family_fixed_set} on the family of functions
    $\varphi_\delta$ with $\delta < \alpha$ with the starting value $0$.  Then
    We have $\varphi_\alpha(0) = \sup\Phi(\alpha)$, and we must show that this
    is equal to $\sup\{\varphi_\delta(0) \mid \delta < \alpha\}$.  We will do so
    by antisymmetry.

    To show that
    \[
        \sup\Phi(\alpha) \leq \sup\{\varphi_\delta(0) \mid \delta < \alpha\},
    \]
    we will use Theorem \ref{ord_normal_family_fixed_least}.  So let $\delta <
    \alpha$.  We must show that
    \[
        \varphi_\delta(\sup\{\varphi_\beta(0) \mid \beta < \alpha\})
        =
        \sup\{\varphi_\beta(0) \mid \beta < \alpha\}.
    \]
    We know that
    \[
        \varphi_\delta(\sup\{\varphi_\beta(0) \mid \beta < \alpha\})
        =
        \sup\{\varphi_\delta(\varphi_\beta(0)) \mid \beta < \alpha\},
    \]
    so it suffices to prove that
    \[
        \sup\{\varphi_\delta(\varphi_\beta(0)) \mid \beta < \alpha\}
        =
        \sup\{\varphi_\beta(0) \mid \beta < \alpha\}.
    \]
    We will prove this by antisymmetry and Theorem \ref{ord_sup_leq_sup}.

    For the forward direction, let $\beta < \alpha$.  We must prove that there
    exists an ordinal $\beta' < \alpha$ such that
    $\varphi_\delta(\varphi_\beta(0)) \leq \varphi_{\beta'}(0)$.  There will be
    two further cases here: when $\delta < \beta$, and when $\delta \geq \beta$.
    When $\delta < \beta$, we have $\varphi_\delta(\varphi_\beta(0)) =
    \varphi_\beta(0)$, so setting $\beta' = \beta$ suffices.  When $\delta \geq
    \beta$, we have $S(\delta) < \alpha$ by Theorem \ref{ord_lim_suc}, and from
    $\beta \leq \delta$ we have $\beta \leq S(\delta)$.  Then, setting $\beta' =
    S(\delta)$, we have
    \[
        \varphi_\delta(\varphi_\beta(0))
        \leq \varphi_\delta(\varphi_{S(\delta)}(0))
        = \varphi_{S(\delta)}(0).
    \]

    For the reverse direction, let $\beta < \delta$.  Then
    \[
        \varphi_\beta(0) \leq \varphi_\delta(\varphi_\beta(0))
    \]
    by Theorem \ref{ord_normal_le}.

    Going all the way back to the original antisymmetry part of the proof, we
    must finally prove
    \[
        \sup\{\varphi_\delta(0) \mid \delta < \alpha\}
        \leq
        \sup\Phi(\alpha).
    \]
    But this one is easy.  This follows directly from Theorem
    \ref{ord_sup_least} and the fact that for any $\delta < \alpha$, we have
    $\varphi_\delta(0) \in \Phi(\alpha)$.
\end{proof}

\begin{theorem} \label{veblen_countable}
    For all countable ordinals $\alpha$ and $\beta$, $\varphi_\alpha(\beta)$ is
    countable.
\end{theorem}
\begin{proof}
    The proof will be by ordinal induction on $\alpha$.  The base case is just
    proving that $\omega^\beta$ is countable, which is Theorem
    \ref{ord_countable_pow}.

    For the successor case, assume that for all countable $\beta$, we have
    $\varphi_\alpha(\beta)$.  We must prove that $\varphi_{S(\alpha)}(\beta) =
    \varphi'_\alpha(\beta)$ is countable.  This follows by Theorem
    \ref{ord_derivative_countable}.

    For the limit case, assume that $\alpha$ is a limit ordinal and that for all
    $\delta < \alpha$ and countable $\varepsilon$, $\varphi_\delta(\varepsilon)$
    is countable as well.  We must prove that $\varphi_\alpha(\beta)$ is
    countable.  We have $\varphi_\alpha(\beta) = \varphi'_{\delta <
    \alpha}(\beta)$, and then the result follows by Theorem
    \ref{ord_family_derivative_countable}.
\end{proof}

\begin{definition}
    Define the Feferman-Sch\"utte ordinal, $\Gamma_0$, to be the first fixed
    point of the normal function $\varphi_\alpha(0)$.  By definition, we know
    that $\varphi_{\Gamma_0}(0) = \Gamma_0$ and that $\Gamma_0$ is the least
    ordinal with this property.
\end{definition}

\begin{theorem}
    The Feferman-Sch\"utte ordinal is countable.
\end{theorem}
\begin{proof}
    Inductively, all ordinals of the form $\varphi_0(0)$,
    $\varphi_{\varphi_0(0)}(0)$, and so on are countable, so $\Gamma_0$ is a
    countable supremum of countable ordinals.
\end{proof}

\end{document}
